---
title: "Brain tumor continuity"
author: "Kim Johnson"
date: `r Sys.Date()`
output: html_document
---

# The main code to generate input files is in "Analysis code 0 to 19.Rmd"
# Open libraries
```{r}
# install.packages("pacman")
pacman::p_load(readr, dplyr, tidyverse, lubridate, survival, survminer, hablar, table1, patchwork, openxlsx, DiagrammeR, lmtest, forestplot, MASS, nnet, broom, forcats, sandwich, mlogit, sjstats, odds.n.ends, labelled, readxl, multcomp, DiagrammeRsvg, rsvg)  
```

# Import files used for analysis
```{r}
seer0to39<-read_csv("/Volumes/kijohnson/Active/seer0to39.csv", show_col_types = FALSE) # this is 0 to 39 cancer patient data

enroll_seerids<-read_csv("/Volumes/kijohnson/Active/enroll_seerids.csv", show_col_types = FALSE) # this is 0 to 39 enrollee data

mcaid <- read_xlsx("/Volumes/kijohnson/Active/mcaid0to39KJ.xlsx") # created in continuity coding file

```

# Identify ids to exclude patients with Medicaid files in multiple states (this is used below)
```{r}
df <- enroll_seerids %>% 
   find_duplicates(patient_id, YEAR) # this identifies patients in multiple states in the same year
 
x <- unique(df$patient_id) # 3588 vector of unique ids who had enrollment in multiple states to exclude in exclusion section

colnames(df)

```

# Reduce cancer dataset to variables of interest for this analysis
```{r}
# select variables
seer0to39<-seer0to39 %>%
  dplyr::select("patient_id", "SEER_registry",  "Race_ethnicity", "sex", "Agerecodewithsingleages_and_100",
"Sequence_number", "Month_of_diagnosis", "Year_of_diagnosis",
"Month_of_diagnosis_recode", "Primary_Site",  "Histology_ICD_O_2",
"Behavior_code_ICD_O_2", "Histologic_Type_ICD_O_3", "Behavior_code_ICD_O_3",
"Grade", "Diagnostic_Confirmation", "Type_of_Reporting_Source", 
"SEERCombinedSummaryStage2000200", 
"Age_recode_with_1_year_olds", "Site_recode_ICD_O_3_WHO_2008",
"ICCCsiterecode_ICD_O_3_WHO_2008", "ICCCsiterecextendedICDO3WHO2008",
"Behavior_recode_for_analysis", "Histologyrecode_broad_groupings", "Race_recode_White_Black_Other",
"Race_recode_W_B_AI_API", "OriginrecodeNHIAHispanicNonHisp",
"Firstmalignantprimary_indicator",
"state", "county", "Vitalstatusrecodestudycutoffuse", 
"AYA_site_recode_WHO_2008", 
"SEERcausespecificdeathclassific", "SEERothercauseofdeathclassifica", "Census_Tract_1990", "Census_Tract_2000", "Census_Tract_2010", "Census_Coding_System", "Census_Tract_Certainty_1990", "Census_Tract_Certainty_2000",
"Census_Tract_Certainty_2010", "Census_Tract_Poverty_Indicator",
"Rural_Urban_Continuum_Code_1993", "Rural_Urban_Continuum_Code_2003",
"Rural_Urban_Continuum_Code_2013", "Health_Service_Area", "HealthService_Area_NCI_Modified",
"SEER_DateofDeath_Month", "SEER_DateofDeath_Year", "Month_of_last_follow_up_recode", "Year_of_last_follow_up_recode", "Year_of_birth", "Date_of_diagnosis_flag", "Date_therapy_started_flag",
"Date_of_birth_flag", "Date_of_last_follow_up_flag", "Month_therapy_started",
"Year_therapy_started",  "Derived_AJCC_flag","Primary_Payer_at_DX",
"COD_ICD_code", "Cause_of_Death_ICD_8_or_9", "Cause_of_Death_ICD_10",
"Recode_ICD_0_2_to_9", "Recode_ICD_0_2_to_10", "NHIA_Derived_Hisp_Origin",
"Insurance_Recode_2007", "Yost_ACS_2006_2010", "Yost_ACS_2010_2014", "Yost_ACS_2013_2017", "Yost_ACS_2006_2010_State_based", "Yost_ACS_2010_2014_State_based", "Yost_ACS_2013_2017_State_based",
"Yost_ACS_2006_2010_quintile", "Yost_ACS_2010_2014_quintile",
"Yost_ACS_2013_2017_quintile", "YostACS20062010quintileStatebas",
"YostACS20102014quintileStatebas", "YostACS20132017quintileStatebas",
"HPV_recode_2010", "In_Medicaid_Flag", "CS_tumor_size_2004_2015", "RX_Summ_Surg_Prim_Site_1998", "Reasonnocancer_directed_surgery", "Radiation_recode", "Chemotherapy_recode_yes_no_unk", "RX_Summ_Systemic_Surg_Seq", "RX_Summ_Surg_Rad_Seq", "Radiation")

table(seer0to39$county)
```

# Data management of variables 
```{r}
# rename age variable to age
seer0to39 <- seer0to39 %>%
  rename(age = Agerecodewithsingleages_and_100)

# recode demographic variables
# Age, race, Hispanic, sex  cause of death, cause specific death

seer0to39<-seer0to39 %>%
  mutate(age_cat=case_when(Age_recode_with_1_year_olds %in% c(0,1) ~ 0, # 0 to 4
                                Age_recode_with_1_year_olds == 2 ~1, # 5 to 9
                                Age_recode_with_1_year_olds == 3 ~2, # 10 to 14
                                Age_recode_with_1_year_olds == 4 ~3, # 15 to 19
                                Age_recode_with_1_year_olds %in% c(5,6,7,8) ~ 4)) %>% # 20 to 39
         mutate(race_cat=case_when(Race_recode_W_B_AI_API == 1 ~0, # White
                                Race_recode_W_B_AI_API == 2 ~1, # Black
                                Race_recode_W_B_AI_API == 3 ~2, # AIAN
                                Race_recode_W_B_AI_API == 4 ~3, # API
                                Race_recode_W_B_AI_API == 9 ~4)) %>% # Unknown
        mutate(hispanic = case_when(OriginrecodeNHIAHispanicNonHisp == 0 ~0, # Non-Hispanic
                                    OriginrecodeNHIAHispanicNonHisp == 1 ~1)) %>% # Hispanic
        mutate(race_ethnicity = case_when(race_cat == 0 & hispanic == 0 ~ 0, # Non-Hispanic White
                                race_cat == 1 & hispanic == 0 ~ 1, # Non-Hispanic Black
                                race_cat == 3 & hispanic == 0 ~ 2, # Non-Hispanic API
                                race_cat %in% c(2,4) & hispanic == 0 ~ 3, # Non-Hispanic Other
                                race_cat %in% c(0:4) & hispanic == 1 ~ 4)) %>% # Hispanic
        mutate(race_cat2 = case_when(race_cat == 0 ~ 0, # Non-Hispanic White
                                     race_cat == 1 ~ 1, # Non-Hispanic Black
                                     race_cat %in% c(2,3,4,5) ~ 2)) %>% # Non-Hispanic Other 
        mutate(race_ethnicity2 = if_else(race_ethnicity %in% c(2,3), 2, 
                                 if_else(race_ethnicity == 4, 3, race_ethnicity))) %>%
        mutate(sex = case_when(sex == 1 ~ 0, # "Male"
                               sex == 2 ~ 1)) %>% # "Female"
        mutate(can_dead = case_when(SEERcausespecificdeathclassific == 0 ~ 0, # Alive or dead of other cause
                                SEERcausespecificdeathclassific == 1 ~ 1)) # Dead from cancer
# note 8 an 9 are missing or N/A
seer0to39$age_cat<-factor(seer0to39$age_cat, levels=c(0:4), labels=c("0-4", "5-9", "10-14", "15-19", "20-39"))

# create age_cat 2 variable for groups 0 to 19 and 20 to 39
seer0to39 <- seer0to39 %>%
  mutate(age_cat2 = case_when(age_cat != "20-39" ~ 0,
                              age_cat == "20-39" ~ 1))

seer0to39$age_cat2 <- factor(seer0to39$age_cat2, levels = c(0:1), labels = c("0 to 19", "20 to 39"))

table(seer0to39$age_cat, seer0to39$age_cat2)

# create 0 to 14 and 15 to 39 age categories
seer0to39 <- seer0to39 %>%
  mutate(age_cat3 = case_when(age_cat %in% c("0-4", "5-9", "10-14") ~ 0,
                              age_cat %in% c("15-19", "20-39") ~ 1))

# Factor variables
seer0to39$age_cat3 <- factor(seer0to39$age_cat3, levels = c(0:1), labels = c("0 to 14", "15 to 39"))
table(seer0to39$age_cat, seer0to39$age_cat3)

seer0to39$race_cat<-factor(seer0to39$race_cat, levels=c(0:4), labels=c("White", "Black", "AIAN", "API", "Unknown"))
seer0to39$hispanic<-factor(seer0to39$hispanic, levels=c(0:1), labels=c("Non-Hispanic", "Hispanic"))
seer0to39$race_ethnicity<-factor(seer0to39$race_ethnicity, levels=c(0:4), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other",  "Hispanic")) 
seer0to39$race_ethnicity2<-factor(seer0to39$race_ethnicity2, levels=c(0:3), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API/Other", "Hispanic")) 
seer0to39$sex <- factor(seer0to39$sex, levels = c(0:1), labels = c("Male", "Female"))
seer0to39$race_cat2<-factor(seer0to39$race_cat2, levels=c(0:2), labels=c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic Other"))

# check variables
table(seer0to39$SEERcausespecificdeathclassific, useNA = "always")
table(seer0to39$age_cat, useNA = "always")
table(seer0to39$race_cat, useNA = "always")
table(seer0to39$hispanic, useNA = "always")
table(seer0to39$race_ethnicity, useNA = "always")
table(seer0to39$can_dead, useNA = "always")
table(seer0to39$sex, useNA = "always")

# calculate follow-up time from month of dx, recode year of dx TO Month_of_last_follow_up_recode, Year_of_last_follow_up_recode. The difference between these variable dates would provide months of follow-up
summary(seer0to39$Month_of_diagnosis_recode)
summary(seer0to39$Year_of_diagnosis)
summary(as.numeric(seer0to39$Month_of_last_follow_up_recode))
seer0to39$Month_of_last_follow_up_recode<-as.numeric(seer0to39$Month_of_last_follow_up_recode) # make variable numeric
seer0to39$SEER_DateofDeath_Month<-as.numeric(seer0to39$SEER_DateofDeath_Month) # make variable numeric
summary(seer0to39$Month_of_last_follow_up_recode)
summary(seer0to39$Year_of_last_follow_up_recode)
summary(seer0to39$SEER_DateofDeath_Month)
summary(seer0to39$SEER_DateofDeath_Year)

# assign a date of diagnosis using 15 as the day because there is no day of dx so assume mean is 15
seer0to39$proxy_day_of_dx <-15
seer0to39$proxy_day_of_fu <-15
seer0to39$proxy_day_of_dth <-15

seer0to39<-seer0to39 %>%
  mutate(proxy_dx_date = make_date(Year_of_diagnosis, Month_of_diagnosis_recode, proxy_day_of_dx)) %>% # make date of dx
  mutate(proxy_fu_date = make_date(Year_of_last_follow_up_recode, Month_of_last_follow_up_recode, proxy_day_of_fu)) %>% # make date of follow-up
  mutate(proxy_dth_date = make_date(SEER_DateofDeath_Year, SEER_DateofDeath_Month, proxy_day_of_dth)) # make date of death

# calculate the difference in days between the estimated date of dx and estimated date of last follow-up for non-cases and cases (actually does not matter if you separate these out or not)

test <- seer0to39 %>% filter (can_dead ==1) 
test$result <- test$proxy_fu_date - test$proxy_dth_date
summary(as.numeric(test$result))
  
seer0to39$fudays<-as.numeric(difftime(seer0to39$proxy_fu_date, seer0to39$proxy_dx_date, units=c("days")))

# estimate follow-up months by dividing by the average number of days in a month
seer0to39$fumo<-seer0to39$fudays/30.437

summary(seer0to39$fumo)

# cancer type coding
table(seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008)

# ICCC3 (https://seer.cancer.gov/iccc/iccc-who2008.html)
seer0to39<-seer0to39 %>%
  mutate(can_type = case_when(
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(11, 12,13,14,15)~0,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(21,22,23,24,25)~1,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(31,32,33,34,35,36)~2,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(41,42) ~3,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(50) ~4,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(61,62,63) ~5,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(71,72,73) ~6,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(81,82, 83,84,85) ~7,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(91,92,93,94,95)~8,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(101,102,103, 104,105) ~9,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(111,112,113,114,115,116) ~10,
    ICCCsiterecode_ICD_O_3_WHO_2008 %in% c(121,122)~11))

seer0to39$can_type<-factor(seer0to39$can_type, levels=c(0:11), labels=c(
  "I. Leukemias, myeloproliferative diseases, and myelodysplastic diseases",
  "II. Lymphomas and reticuloendothelial neoplasms",
  "III. CNS and miscellaneous intracranial and intraspinal neoplasms",
  "IV. Neuroblastoma and other peripheral nervous cell tumors",
  "V. Retinoblastoma",
  "VI. Renal tumors",
  "VII. Hepatic tumors",
  "VIII. Malignant bone tumors",
  "IX. Soft tissue and other extraosseous sarcomas",
  "X. Germ cell tumors, trophoblastic tumors, and neoplasms of gonads",
  "XI. Other malignant epithelial neoplasms and malignant melanomas",
  "XII. Other and unspecified malignant neoplasms"))
    
table(seer0to39$can_type, seer0to39$ICCCsiterecode_ICD_O_3_WHO_2008, useNA= "always")  

table(seer0to39$AYA_site_recode_WHO_2008)

# use Yost index as a measure of poverty
table(seer0to39$Yost_ACS_2006_2010_quintile,  useNA = "always")

class(seer0to39$Yost_ACS_2006_2010_quintile)

seer0to39$Yost_ACS_2006_2010_quintile <-factor(seer0to39$Yost_ACS_2006_2010_quintile, levels = c(1:5), labels = c(
        "Group 1 (lowest SES)",
        "Group 2",
        "Group 3",
        "Group 4",
        "Group 5 (highest SES)"))

# classify stage
table(seer0to39$SEERCombinedSummaryStage2000200, useNA = "always")
seer0to39 <- seer0to39 %>%
  mutate(stage = case_when(SEERCombinedSummaryStage2000200 %in% c(0,1) ~ 0,
                           SEERCombinedSummaryStage2000200 %in% c(2,3,4,5) ~ 1,
                           SEERCombinedSummaryStage2000200 %in% c(7) ~ 2))

seer0to39$stage <- factor(seer0to39$stage, levels = c(0:2), labels = c("In situ/localized", "Regional", "Distant"))

table(seer0to39$stage, useNA ="always")

table(seer0to39$SEER_DateofDeath_Year, useNA = "always")

table(seer0to39$can_type)

# create categories for AYA codes
class(seer0to39$AYA_site_recode_WHO_2008) # numeric

seer0to39 <- seer0to39 %>%
  mutate(can_typeAYA = case_when(AYA_site_recode_WHO_2008 %in% c(1,2,3,4) ~ 0,
                                 AYA_site_recode_WHO_2008 %in% c(5,6) ~ 1,
                                 AYA_site_recode_WHO_2008 %in% c(7,8,9,10,11,12,13,14,15,16) ~ 2,
                                 AYA_site_recode_WHO_2008 %in% c(17,18,19,20) ~ 3,
                                 AYA_site_recode_WHO_2008 %in% c(21,22,23,24,25) ~ 4,
                                 AYA_site_recode_WHO_2008 %in% c(26,27,28) ~ 5,
                                 AYA_site_recode_WHO_2008 %in% c(29,30) ~ 6,
                                 AYA_site_recode_WHO_2008 %in% c(31,32,33,34,35,36,37,38,39,40,41,42,43,44,45,46,47,48) ~ 7,
                                 AYA_site_recode_WHO_2008 %in% c(49,50,51,52,53,54,55) ~ 8,
                                 AYA_site_recode_WHO_2008 %in% c(56) ~ 9,
                                 AYA_site_recode_WHO_2008 == 99 ~ 10))
                                 
seer0to39$can_typeAYA <- factor(seer0to39$can_typeAYA, levels = c(0:10), labels = c(
  "1. Leukemias",
  "2. Lymphomas",
  "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)",
  "4. Osseous & Chondromatous Neoplasms",
  "5. Soft Tissue Sarcomas",
  "6. Germ Cell and Trophoblastic Neoplasms",
  "7. Melanoma and Skin Carcinomas",
  "8. Carcinomas",
  "9. Miscellaneous specified neoplasms, NOS",
  "10. Unspecified Malignant Neoplasms",
  "Unclassified"))
table(seer0to39$can_typeAYA, useNA = "always")


# categorize brain tumors according to AYA definition

seer0to39 <- seer0to39 %>%
  mutate(brain = case_when (can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 %in% c(7,8,9) ~ 0, # 3.1 Astrocytoma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(10) ~ 1, # 3.2 Other glioma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(11) ~ 2, # 3.3 Ependymoma
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(12,13) ~ 3, # 3.4 Medulloblastoma and other PNET
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(14) ~ 4, # 3.5 Other specified intracranial and instraspinal neoplasms
                            can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)" & AYA_site_recode_WHO_2008 %in% c(15,16) ~ 5)) # 3.6 Unspecified intracranial and intraspinal neoplasms
seer0to39$brain
table(seer0to39$brain, useNA = "always")

# make brain2 for Table 1 because of small numbers
# combine because of small numbers (Non-Hispanic API and Non-Hispanic  Other, 3.5 Other specified intracranial and instraspinal neoplasms, 3.6 Unspecified intracranial and intraspinal neoplasms)
seer0to39 <- seer0to39 %>%
  mutate(brain2 = if_else(brain %in% c(4,5), 4, brain)) 

table(seer0to39$brain2)

seer0to39$brain <- factor(seer0to39$brain, levels = c(0:5), labels = c(
  "3.1 Astrocytoma",
  "3.2 Other glioma",
  "3.3 Ependymoma",
  "3.4 Medulloblastoma and other PNET",
  "3.5 Other specified intracranial and instraspinal neoplasms",
  "3.6 Unspecified intracranial and intraspinal neoplasms"))
table(seer0to39$brain, useNA = "always")


seer0to39$brain2 <- factor(seer0to39$brain2, levels = c(0:4), labels = c(
  "3.1 Astrocytoma",
  "3.2 Other glioma",
  "3.3 Ependymoma",
  "3.4 Medulloblastoma and other PNET",
  "Other specified (3.5) and unspecified (3.6) intracranial and instraspinal neoplasms"))
table(seer0to39$brain2, useNA = "always")

# further classify low grade astrocytomas for sensitivity analysis (exclude low grade)
seer0to39 <- seer0to39 %>%
  mutate(lowgrade_exclude = case_when(can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 == 7 ~ 1, # these are low grade tumors
                              can_typeAYA == "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)"  & AYA_site_recode_WHO_2008 != 7 ~ 0)) # those are not low grade tumors


table(seer0to39$brain, seer0to39$Grade)

# recode treatment variables

# Surgery and reason
seer0to39 <- seer0to39 %>%
  mutate(surgery = case_when(RX_Summ_Surg_Prim_Site_1998 == 0 ~ 0, # no surgery
                             RX_Summ_Surg_Prim_Site_1998 >0 & RX_Summ_Surg_Prim_Site_1998 <99 ~ 1, # surgery
                             RX_Summ_Surg_Prim_Site_1998 == 99 ~ 2)) %>% # unknown surgery
  mutate(surg_reason = case_when(Reasonnocancer_directed_surgery == 0 ~ 0, # surgery performed
                                 Reasonnocancer_directed_surgery == 1 ~ 1, # surgery not recommended
                                 Reasonnocancer_directed_surgery == 2 ~ 2, # contraindicated/autopsy only
                                 Reasonnocancer_directed_surgery == 5 ~ 3, # patient died before recommended surgery
                                 Reasonnocancer_directed_surgery == 6 ~ 4, # unknown reason for no surgery
                                 Reasonnocancer_directed_surgery == 7 ~ 5, # patient or patient's guardian refused
                                 Reasonnocancer_directed_surgery == 8 ~ 6, # recommended, unknown if done
                                 Reasonnocancer_directed_surgery == 9 ~ 7)) # unknown, autopsy or death certificate only case

seer0to39$surgery <- factor(seer0to39$surgery, levels = c(0:2), labels = c("No surgery", "Surgery", "Unknown"))
  
seer0to39$surg_reason <- factor(seer0to39$surg_reason, levels = c(0:7), labels = c("Surgery performed", "Surgery not recommended", "Contraindicated/autopsy only", "Patient died before recommended surgery", "Unknown reason for no surgery", "Patient or patient's guardian refused", "Recommended, unknown if done", "Unknown, autopsy or death certificate only case"))
        

table(seer0to39$surgery)
table(seer0to39$surg_reason, useNA = "always")

# Radiation
table(seer0to39$Radiation_recode, useNA = "always")

seer0to39 <- seer0to39 %>%
  mutate(radiation = case_when(Radiation %in% c(0,7) ~ 0, # none, diagnosed at autopsy, unknown
                             Radiation %in% c(1,2,3,4,5,6) ~ 1, # radiation
                             Radiation %in% c(8,9) ~ 2)) # unknown

table(seer0to39$radiation, seer0to39$Radiation, useNA="always")
seer0to39$radiation<- factor(seer0to39$radiation, levels = c(0:2), labels = c("No/unknown radiation", "Radiation", "Unknown"))
  
table(seer0to39$radiation, useNA = "always")

# Chemotherapy
table(seer0to39$Chemotherapy_recode_yes_no_unk) # not provided

# code surgery and radiation together
seer0to39 <- seer0to39 %>%
  mutate(surgery_radiation = case_when(surgery == "No surgery" & radiation == "No/unknown radiation" ~0,
                                       surgery == "No surgery" & radiation == "Radiation" ~ 1,
                                       surgery == "No surgery" & radiation == "Unknown" ~ 2,
                                       surgery == "Surgery" & radiation == "No/unknown radiation" ~3,
                                       surgery == "Surgery" & radiation == "Radiation" ~ 4,
                                       surgery == "Surgery" & radiation == "Unknown" ~ 5,
                                       surgery == "Unknown" & radiation == "No/unknown radiation" ~6,
                                       surgery == "Unknown" & radiation == "Radiation" ~ 7,
                                       surgery == "Unknown" & radiation == "Unknown" ~ 8))

seer0to39$surgery_radiation <- factor(seer0to39$surgery_radiation, levels = c(0:8), labels = c("No surgery, No/unknown radiation", "No surgery, radiation", "No surgery, unknown",
                                                                                               "Surgery, No/unknown radiation", "Surgery, radiation", "Surgery, unknown",
                                                                                               "Unknwon, No/unknown radiation", "Unknown, radiation", "Unknown, unknown"))

# code surgery and radiation together
seer0to39 <- seer0to39 %>%
  mutate(surgery_radiation2 = case_when(surgery == "No surgery" & radiation %in% c("No/unknown radiation", "Unknown") ~ 0,
                                       surgery %in% c("No surgery", "Unknown") & radiation %in% c("Radiation") ~ 1,
                                       surgery == "Surgery" & radiation == "Radiation" ~ 2,
                                       surgery == "Surgery" & radiation %in% c("Unknown", "No/unknown radiation") ~ 3,
                                       surgery == "Unknown" & radiation %in% c("No/unknown radiation", "Unknown") ~ 4))

seer0to39$surgery_radiation2 <- factor(seer0to39$surgery_radiation2, levels = c(0:4), labels = c("No surgery, No/unknown radiation", "No surgery/surgery unknown, radiation", "Surgery and Radiation", "Surgery and No/unknown radiation", "Unknown surgery and no/unknown radiation"))

table(seer0to39$surgery_radiation, useNA = "always")
table(seer0to39$surgery_radiation, seer0to39$surgery_radiation2, useNA = "always")
```

# Exclusions

# limit to brain tumors
```{r}
 seer0to39 <- seer0to39 %>%
  filter(can_typeAYA== "3. CNS and Other Intracranial and Intraspinal Neoplasms (all behaviors)") # 11274

table(seer0to39$brain)

```
# limit to first cancers
```{r}
seer0to39 <- seer0to39 %>%
  filter(Firstmalignantprimary_indicator == 1) # 10,724
```

# limit to sequence number 01 or 00
```{r}
seer0to39 <- seer0to39 %>%
  filter(Sequence_number %in% c("00", "01")) # 10,721
```

# Exclude multiple state Medicaid ids within a single year defined above
```{r}
seer0to39 <- seer0to39 %>% 
  filter(!patient_id %in% x) # 10,449
```

# Exclude those with Medicaid cancer registry match scores below  5
```{r}
match <- enroll_seerids %>%
  filter(MATCH_SCORE_HI <5)

# create list of unique ids to exclude
ids_2_exclude <- unique(match$patient_id)


seer0to39 <- seer0to39 %>%
  filter(!patient_id %in% ids_2_exclude) # 10,398
```

# Exclude those with missing data on variables used in the analysis
```{r}
seer0to39<-seer0to39 %>%
  filter(!is.na(fumo)) # 10,398

seer0to39<-seer0to39 %>%
  filter(!is.na(can_dead)) # 10,302

seer0to39<-seer0to39 %>%
  filter(!is.na(In_Medicaid_Flag)) # 10,302

seer0to39<-seer0to39 %>%
  filter(!is.na(race_ethnicity)) # 10,302

seer0to39<-seer0to39 %>%
  filter(!is.na(age_cat)) # 10,302

table(seer0to39$Yost_ACS_2006_2010_quintile, useNA = "always")
seer0to39<-seer0to39 %>%
  filter(!is.na(Yost_ACS_2006_2010_quintile)) # 10,107

missing_data_percent  <- 100- (10107/10398)*100
missing_data_percent  # <2.8%

```

# merge in mcont (medicaid enrollment data and recode 0 = Not linked, 1 = continuous, 2 = discontinuous, 3 = Other)
```{r}
seer0to39b <- seer0to39
seer0to39 <- left_join(seer0to39b, mcaid, by = "patient_id")

table(seer0to39$mcontb, useNA = "always")
table(mcaid$mcontb, useNA = "always")
table(seer0to39$In_Medicaid_Flag)

# recode a category for not in medicaid
seer0to39 <- seer0to39 %>%
  mutate(mcont66 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb == 0 ~ 1, # continuous
                             mcontb == 1 ~ 2, # discontinuous
                             mcontb == 2 ~ 3)) # Other
table(seer0to39$mcont66, useNA = "always")

table(seer0to39$mcont66, seer0to39$In_Medicaid_Flag, useNA = "always") 


# create factor variable for analyses
seer0to39$mcont66 <- factor(seer0to39$mcont66, levels = c(0:3), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other"))

table(seer0to39$mcont66, seer0to39$In_Medicaid_Flag, useNA = "always") 

# for definition 2

# recode a category for not in medicaid
seer0to39 <- seer0to39 %>%
  mutate(mcont66.2 = case_when(In_Medicaid_Flag == 0 ~ 0, # not linked
                             mcontb2 == 0 ~ 1, # continuous
                             mcontb2 == 1 ~ 2, # discontinuous
                             mcontb2 == 2 ~ 3, # other
                             mcontb2 == 3 ~ 4)) # at diagnosis


# create factor variable for analyses
seer0to39$mcont66.2 <- factor(seer0to39$mcont66.2, levels = c(0:4), labels = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"))

table(seer0to39$mcont66.2, useNA = "always")
```


# Figure 1
```{r}
Figure <- grViz("digraph flowchart {
  # node definitions with substituted label text
  node [fontname = Helvetica, shape = rectangle, fontsize=24, fontcolor='blue']
  tab1 [label = '@@1']
  tab2 [label = '@@2']
  tab3 [label = '@@3']
  tab4 [label = '@@4']
  tab5 [label = '@@5']
  tab6 [label = '@@6']
  tab7 [label = '@@7']
  # edge definitions with the node IDs
  tab1 -> tab2 -> tab3 -> tab4 -> tab5 -> tab6 -> tab7;
}
[1]: 'Records received for individuals diagnosed with brain cancers at ages 0 to 39 years from 2006-2013 n = 11,274'
[2]: 'Excluding 550 individuals whose first malignant primary indicator was not equal to 1 n = 10,724'
[3]: 'Excluding 3 individuals whose cancer sequence was > 1 n = 10,721'
[4]: 'Excluding 272 individuals linked to multiple state Medicaid files n = 10,449'
[5]: 'Excluding 51 individuals with Medicaid enrollment match scores <5 n =  10,398'
[6]: 'Excluding 96 individuals with missing data on cancer specific cause of death n = 10,302'
[7]: 'Excluding 195 individuals with missing data on Yost ACS 2006 to 2010 Quintile n = 10,107'
")

Figure


Figure %>%
   DiagrammeRsvg::export_svg() %>% 
  charToRaw() %>% 
  rsvg::rsvg_pdf("Figure 1.pdf")
```

# TABLE 1
```{r}
label(seer0to39$age_cat)<-"Age category"
label(seer0to39$sex)<-"Sex"
label(seer0to39$race_ethnicity)<-"Race/ethnicity category"
label(seer0to39$fumo)<-"Follow-up months"
label(seer0to39$can_typeAYA)<-"Cancer type"
label(seer0to39$Yost_ACS_2006_2010_quintile) <- "Yost Index 2006 to 2010"
label(seer0to39$stage) <- "Stage"
label(seer0to39$brain) <- "Brain tumor type"
label(seer0to39$mcont66) <- "Medicaid continuity"
label(seer0to39$brain2) <- "Brain tumor type"
label(seer0to39$race_ethnicity2) <-"Race/ethnicity category"
label(seer0to39$mcont66.2) <- "Medicaid continuity"
label(seer0to39$radiation) <- "Radiation"
label(seer0to39$surgery) <- "Surgery"
label(seer0to39$surgery_radiation2) <- "Surgery and radiation"

# make dead a factor variable just for this table
seer0to39$can_dead2<-factor(seer0to39$can_dead, levels=c(0,1), labels= c("Alive", "Deceased"))
table(seer0to39$can_dead2)
label(seer0to39$can_dead2) <- "Vital Status"
seer0to39$In_Medicaid_Flag<-factor(seer0to39$In_Medicaid_Flag, levels=c(0,1), labels= c("Not Enrolled", "Enrolled"))
table(seer0to39$can_dead2)
table(seer0to39$In_Medicaid_Flag)

library(table1)

# Table 1 for paper
table1(~ sex + race_ethnicity + mcont66 + Yost_ACS_2006_2010_quintile + brain2 + stage + surgery_radiation2 + can_dead2|age_cat3 + In_Medicaid_Flag, droplevels = TRUE, seer0to39)

```


# Stage at DX analysis
```{r}
seer0to39_stage <- seer0to39 %>%
  filter(!is.na(stage)) %>% # this yields 
  select(stage, In_Medicaid_Flag, SEER_registry, mcont66.2, race_ethnicity2, race_cat2, age_cat, age, Yost_ACS_2006_2010_quintile, state, can_type, brain, brain2, age_cat2, age_cat3) %>%
  mutate(stage2 = case_when(stage == "In situ/localized" ~ 0, 
                            stage %in% c("Regional", "Distant") ~ 1)) # combine regional and distant

seer0to39_stage$stage2 <- factor(seer0to39_stage$stage2, levels = c(0,1), labels = c("In situ/localized", "Regional/Distant"))

table(seer0to39_stage$stage2, seer0to39_stage$brain)
class(seer0to39_stage$race_cat2)

# In_Medicaid_Flag adjusted model 0 to 14

model1 <- miceadds::glm.cluster(seer0to39_stage %>% filter(age_cat3 == "0 to 14"), stage2 ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model1)

# extract coefficients 
b <-as.data.frame(exp(coef(model1)))
# CI(95%)
CI<- as.data.frame(exp(confint(model1, level=0.95)))
SE <- as.data.frame(sqrt(diag(vcov(model1))))
p <- (1-pnorm(abs(coef(model1)/sqrt(diag(vcov(model1)))))) * 2
df <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(df)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
df <- tibble::rownames_to_column(df, "Exposure comparison") # Apply rownames_to_column
df<-df[-c(1,3:13),]

# Replace multiple strings at a time
rep_str = c('In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
df$`Exposure comparison`<- str_replace_all(df$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df$`Age group` <- "0 to 14"

# In_Medicaid_Flag adjusted model 15 to 39
model2 <- miceadds::glm.cluster(seer0to39_stage %>% filter(age_cat3 == "15 to 39"), stage2 ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model2)

# extract coefficients  
b <-as.data.frame(exp(coef(model2)))
# CI(95%)
CI<- as.data.frame(exp(confint(model2, level=0.95)))
SE <- as.data.frame(sqrt(diag(vcov(model2))))
p <- (1-pnorm(abs(coef(model2)/sqrt(diag(vcov(model2)))))) * 2
df2 <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(df2)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
df2 <- tibble::rownames_to_column(df2, "Exposure comparison") # Apply rownames_to_column
df2<-df2[-c(1,3:13),]

# Replace multiple strings at a time
rep_str = c('In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
df2$`Exposure comparison`<- str_replace_all(df2$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df2$`Age group` <- "15 to 39"
```

# mcont66.2
```{r}
# mcont.66 adjusted model 0 to 19
model1 <- miceadds::glm.cluster(seer0to39_stage %>% filter(age_cat3 == "0 to 14"), stage2 ~ mcont66.2 + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model1)

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
b <-as.data.frame(exp(coef(model1)))
CI<- as.data.frame(exp(confint(model1, level=0.95)))
SE <- as.data.frame(sqrt(diag(vcov(model1))))
p <- (1-pnorm(abs(coef(model1)/sqrt(diag(vcov(model1)))))) * 2
df3 <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(df3)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
df3 <- tibble::rownames_to_column(df3, "Exposure comparison") # Apply rownames_to_column

df3<-df3[-c(1,6:16),]

# Replace multiple strings at a time
rep_str = c('mcont66.2Continuous'='Continuous vs. Not enrolled', 'mcont66.2Discontinuous'='Discontinuous vs. Not enrolled',
  'mcont66.2Other'='Other vs. Not enrolled', 'mcont66.2At diagnosis'='At diagnosis vs. Not enrolled')
df3$`Exposure comparison`<- str_replace_all(df3$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df3$`Age group` <- "0 to 14"

# In_Medicaid_Flag adjusted model 15 to 39
model2 <- miceadds::glm.cluster(seer0to39_stage %>% filter(age_cat3 == "15 to 39"), stage2 ~ mcont66.2 + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model2)

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
b <-as.data.frame(exp(coef(model2)))
CI<- as.data.frame(exp(confint(model2, level=0.95)))
SE <- as.data.frame(sqrt(diag(vcov(model2))))
p <- (1-pnorm(abs(coef(model2)/sqrt(diag(vcov(model2)))))) * 2
df4 <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(df4)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
df4 <- tibble::rownames_to_column(df4, "Exposure comparison") # Apply rownames_to_column

df4 <- df4[-c(1,6:16),]

# Replace multiple strings at a time
rep_str = c('mcont66.2Continuous'='Continuous vs. Not enrolled', 'mcont66.2Discontinuous'='Discontinuous vs. Not enrolled',
  'mcont66.2Other'='Other vs. Not enrolled', 'mcont66.2At diagnosis'='At diagnosis vs. Not enrolled')
df4$`Exposure comparison`<- str_replace_all(df4$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df4$`Age group` <- "15 to 39"
```

# Table 2: Stage at diagnosis analysis
```{r}
Table2 <- rbind(df,df2,df3,df4)
Table2
```

# Plots of stage at dx (NO FIGURE, just a visual for easy interpretation)
```{r}
Plot1 <- Table2 %>%
  filter(`Exposure comparison` == "Enrolled vs. Not enrolled") %>%
  ggplot() +
  geom_point(aes(y=OR, x = `Age group`, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=`Age group`, group = `Exposure comparison`, color = `Exposure comparison`, ymax = highCI, ymin = lowCI), width = 0.1,  color = "black") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "A. Enrolled vs. Not enrolled", x = "Age group") +
  ylim(0.75,1.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black")) 

Plot1


Plot2 <- Table2 %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled") 
Plot2$`Exposure comparison` <- factor(Plot2$`Exposure comparison`, levels = c("At diagnosis vs. Not enrolled", "Discontinuous vs. Not enrolled",  "Continuous vs. Not enrolled",  "Other vs. Not enrolled"))

Plot2 <- Plot2 %>%
  ggplot() +
  geom_point(aes(y=OR, x = `Age group`, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x= `Age group`, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = highCI, ymin = lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "B.", x = "Age group") +
  #ylim(0.5,2.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Continuous vs. Not enrolled", "Discontinuous vs. Not enrolled",  "Other vs. Not enrolled", "At diagnosis vs. Not enrolled"), values=c("cyan3", "blue", "black", "red")) + guides(colour = guide_legend(nrow = 2))

Plot2
```
# Stage at dx plots by main cancer types 
```{r}
table(seer0to39_stage$brain2)
limited <- seer0to39_stage %>%
 filter(brain2 %in% c("3.1 Astrocytoma", "3.2 Other glioma",  "3.3 Ependymoma", "3.4 Medulloblastoma and other PNET"))
table(limited$brain2)
ct <- unique(limited$brain2)

# In_Medicaid_Flag 0 to 14
df1 <- data.frame(matrix(ncol=6, nrow=0)) # create the table
columns<-c("Category", "OR","lowCI", "highCI", "BrainType") # label the columns with names
colnames(df1)<-columns

for (i in ct) {
dfa<- limited %>% 
    filter(brain2 == i)
  x <- miceadds::glm.cluster(dfa %>% filter(age_cat3 == "0 to 14"), stage2 ~ In_Medicaid_Flag + race_cat2 + age  + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
b <-as.data.frame(exp(coef(x)))
# CI(95%)
CI<- as.data.frame(exp(confint(x, level=0.95)))
p <- (1-pnorm(abs(coef(x)/sqrt(diag(vcov(x)))))) * 2
dfa <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(dfa)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
dfa <- tibble::rownames_to_column(dfa, "Exposure comparison") # Apply rownames_to_column
dfa<-dfa[-c(1,3:13),]
  
dfa$BrainType<-i # add cancer type
df1 <-rbind(dfa, df1)
}

df1 <- df1[order(df1$BrainType),] # In_Medicaid_Flag

# Replace multiple strings at a time
rep_str = c('In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
df1$`Exposure comparison`<- str_replace_all(df1$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df1$`Age group` <- "0 to 14"


# In_Medicaid_Flag 20 to 39
df2 <- data.frame(matrix(ncol=6, nrow=0)) # create the table
columns<-c("Category", "OR","lowCI", "highCI", "BrainType") # label the columns with names
colnames(df2)<-columns

for (i in ct) {
dfa<- limited %>% 
    filter(brain2 == i)
  x <- miceadds::glm.cluster(dfa %>% filter(age_cat3 == "15 to 39"), stage2 ~ In_Medicaid_Flag + race_cat2 + age  + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
b <-as.data.frame(exp(coef(x)))
# CI(95%)
CI<- as.data.frame(exp(confint(x, level=0.95)))
p <- (1-pnorm(abs(coef(x)/sqrt(diag(vcov(x)))))) * 2
dfa <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(dfa)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
dfa <- tibble::rownames_to_column(dfa, "Exposure comparison") # Apply rownames_to_column
dfa<-dfa[-c(1,3:13),]
  
dfa$BrainType<-i # add cancer type
df2 <-rbind(dfa, df2)
}

df2<- df2[order(df2$BrainType),] # In_Medicaid_Flag

# Replace multiple strings at a time
rep_str = c('In_Medicaid_FlagEnrolled'='Enrolled vs. Not enrolled')
df2$`Exposure comparison`<- str_replace_all(df2$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df2$`Age group` <- "15 to 39"

SuppTable1a <- rbind(df1, df2)
```

# mcont66.2
```{r}
# mcont66.2 0 to 19
df3 <- data.frame(matrix(ncol=6, nrow=0)) # create the table
columns<-c("Category", "OR","lowCI", "highCI", "BrainType") # label the columns with names
colnames(df3)<-columns

for (i in ct) {
dfa<- limited %>% 
    filter(brain2 == i)
  x <- miceadds::glm.cluster(dfa %>% filter(age_cat3 == "0 to 14"), stage2 ~ mcont66.2 + race_cat2 + age  + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
b <-as.data.frame(exp(coef(x)))
# CI(95%)
CI<- as.data.frame(exp(confint(x, level=0.95)))
p <- (1-pnorm(abs(coef(x)/sqrt(diag(vcov(x)))))) * 2
dfa <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(dfa)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
dfa <- tibble::rownames_to_column(dfa, "Exposure comparison") # Apply rownames_to_column
dfa<-dfa[-c(1,6:16),]
  
dfa$BrainType<-i # add cancer type
df3 <-rbind(dfa, df3)
}

df3<- df3[order(df3$BrainType),] # mcont66.2

# Replace multiple strings at a time
rep_str = c('mcont66.2Continuous' = 'Continuous vs. Not enrolled', 'mcont66.2Discontinuous' = 'Discontinuous vs. Not enrolled',
            'mcont66.2Other' = "Other vs. Not enrolled", 'mcont66.2At diagnosis' = 'At diagnosis vs. Not enrolled')
df3$`Exposure comparison`<- str_replace_all(df3$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df3$`Age group` <- "0 to 14"


# In_Medicaid_Flag 15 to 39
df4 <- data.frame(matrix(ncol=6, nrow=0)) # create the table
columns<-c("Category", "OR","lowCI", "highCI", "BrainType") # label the columns with names
colnames(df4)<-columns

for (i in ct) {
dfa<- limited %>% 
    filter(brain2 == i)
  x <- miceadds::glm.cluster(dfa %>% filter(age_cat3 == "15 to 39"), stage2 ~ mcont66.2 + race_cat2 + age  + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")

# extract coefficients with non-clustered SEs from the model and exponentiate for comparison 
b <-as.data.frame(exp(coef(x)))
# CI(95%)
CI<- as.data.frame(exp(confint(x, level=0.95)))
p <- (1-pnorm(abs(coef(x)/sqrt(diag(vcov(x)))))) * 2
dfa <- cbind(b, CI, p)

# relabel columns and drop intercept rows
names(dfa)[1:4]<-c("OR", "lowCI", "highCI", "p-value")
dfa <- tibble::rownames_to_column(dfa, "Exposure comparison") # Apply rownames_to_column
dfa<-dfa[-c(1,6:16),]
  
dfa$BrainType<-i # add cancer type
df4 <-rbind(dfa, df4)
}

df4 <- df4[order(df4$BrainType),] # In_Medicaid_Flag

# Replace multiple strings at a time
rep_str = c('mcont66.2Continuous' = 'Continuous vs. Not enrolled', 'mcont66.2Discontinuous' = 'Discontinuous vs. Not enrolled',
            'mcont66.2Other' = "Other vs. Not enrolled", 'mcont66.2At diagnosis' = 'At diagnosis vs. Not enrolled')
df4$`Exposure comparison`<- str_replace_all(df4$`Exposure comparison`, rep_str) # this is the dataframe for summary results
df4$`Age group` <- "15 to 39"


SuppTable1b <- rbind(df3, df4)
```

# SUPPLEMENTARY TABLE 1: Stage at diagnosis for main cancer types
```{r}
options(scipen = 999)
SuppTable1 <- rbind(SuppTable1a, df3, df4)

# GET N's
seer0to19 <- seer0to39 %>%
  filter(age <15)

seer20to39 <- seer0to39 %>%
  filter(age >=15)

table(seer0to19$mcont66.2, seer0to19$brain2)

table(seer20to39$mcont66.2, seer20to39$brain2)
```

# SUPPLEMENTARY FIGURE 1: Enrolled vs. Not enrolled Plots by cancer type for four main cancer types
```{r}
# graph ORs and 95% CIs
table(SuppTable1$BrainType)
ForPlotting <- SuppTable1 %>%
  mutate(Abbrev = case_when(BrainType == "3.1 Astrocytoma" ~ "AST",
                            BrainType == "3.2 Other glioma" ~ "OGLIO",
                            BrainType == "3.3 Ependymoma" ~ "EPEN",
                            BrainType == "3.4 Medulloblastoma and other PNET" ~ "MB/PNET",
                            ))

ForPlotting$Abbrev <- factor(ForPlotting$Abbrev, levels = c("AST", "OGLIO", "EPEN", "MB/PNET"))

ForPlotting$Abbrev<- fct_rev(ForPlotting$Abbrev)

Plot1 <- ForPlotting %>%
  filter(`Exposure comparison` == "Enrolled vs. Not enrolled", `Age group` == "0 to 14") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`, ymax = highCI, ymin = lowCI), width = 0.1,  color = "black") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "A. 0 to 14", x = "Brain tumor type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black")) 

Plot1

Plot2 <- ForPlotting %>%
  filter(`Exposure comparison` == "Enrolled vs. Not enrolled", `Age group` == "15 to 39") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), color = "black") + 
  geom_hline(yintercept =1, linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`, ymax = highCI, ymin = lowCI), width = 0.1,  color = "black") +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "B. 15 to 39", x = "Brain tumor type") +
  ylim(0,3.5) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Enrolled"), values=c("black")) 

Plot2

```
# SUPPLEMENTARY FIGURE 1: mcont66.2
```{r}
# graph ORs and 95% CIs
ForPlotting$`Exposure comparison` <- factor(ForPlotting$`Exposure comparison`, levels = c("Discontinuous vs. Not enrolled", "Other vs. Not enrolled", "Continuous vs. Not enrolled", "At diagnosis vs. Not enrolled"), labels = c("Discontinuous", "Other", "Continuous", "At diagnosis"))

Plot3 <- ForPlotting %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled",  `Age group` == "0 to 14") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = highCI, ymin = lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "C. 0 to 14", x = "Brain tumor type") +
  ylim(0,7.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Other", "Continuous", "At diagnosis"),
                        values=c("red", "cyan3", "blue", "black")) 

Plot3

Plot4 <- ForPlotting %>%
  filter(`Exposure comparison` != "Enrolled vs. Not enrolled",  `Age group` == "15 to 39") %>%
  ggplot() +
  geom_point(aes(y=OR, x = Abbrev, group = `Exposure comparison`, color = `Exposure comparison`), na.rm=TRUE, position=position_dodge(width=0.9)) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(position = position_dodge(width=0.9), aes(y=OR, x=Abbrev, group = `Exposure comparison`, color = `Exposure comparison`,  ymax = highCI, ymin = lowCI),  width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=11, face = "bold", color = "black"),
    axis.text.y=element_text(size=11, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, title = "C. 15 to 39", x = "Brain tumor type") +
  ylim(0,7.0) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=9, face="bold")) +
  scale_color_manual(breaks = c("Discontinuous", "Other", "Continuous", "At diagnosis"),
                        values=c("red", "cyan3", "blue", "black")) 

Plot4

table(seer0to19$In_Medicaid_Flag, seer0to19$brain2)
table(seer20to39$In_Medicaid_Flag, seer20to39$brain2)


table(seer0to19$mcont66.2, seer0to19$brain2)
table(seer20to39$mcont66.2, seer20to39$brain2)


table(seer0to19$brain2, seer0to19$stage)
table(seer20to39$brain2, seer20to39$stage)

```

# FIGURE 2: KM Survival plots by age group In_Medicaid Flag
```{r}
# In_Medicaid_flag 0 to 14
plot1 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to39 %>% filter(age_cat3 == "0 to 14")), 
                    size = 0.75,
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct",  
    title = "A. 0 to 14 years",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(8, "bold"),
    legend.labs = c("Not enrolled", "Enrolled"),
    palette = c("blue", "black"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(0, 100),
    censor.size=0.25,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(0,10),
    conf.int = FALSE) 
plot1

# In_Medicaid_flag 15 to 39
plot2 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to39 %>% filter(age_cat3 == "15 to 39")), 
                    size = 0.75,
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct",  
    title = "B. 15 to 39 years",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(8, "bold"),
    legend.labs = c("Not enrolled", "Enrolled"),
    palette = c("blue", "black"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(0, 100),
    censor.size=0.25,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(0,10),
    conf.int = FALSE) 
plot2
```

# FIGURE 2: Survival plots by age group mcont66.2
```{r}
# mcont66 0 to 19
plot3 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66.2, data = seer0to39 %>% filter(age_cat3 == "0 to 14")), 
                    size = 0.75,
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct",  
    title = "C. 0 to 14 years",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(8, "bold"),
    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"),
    palette = c("black", "blue", "maroon2", "cyan3", "red", "green", "purple"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(0, 100),
    censor.size=0.25,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(0,10),
    conf.int = FALSE) 
plot3

# mcont66 20 to 39
plot4 <- ggsurvplot(fit = survfit(Surv(fumo, can_dead) ~ mcont66.2, data = seer0to39 %>% filter(age_cat3 == "15 to 39")), 
                    size = 0.75,
    xlab = "Months from diagnosis", 
    ylab = "Survival probability", 
    fun="pct",  
    title = "D. 15 to 39 years",
    font.title = c(12, "bold"),
    legend.title = "",
    legend = "bottom",
    font.legend=c(8, "bold"),
    legend.labs = c("Not enrolled", "Continuous", "Discontinuous", "Other", "At diagnosis"),
    palette = c("black", "blue", "maroon2", "cyan3", "red", "green", "purple"),
    font.x = c(12, "bold"),
    font.y = c(12, "bold"),
    xlim=c(0, 60),
    break.x.by = 12,
    ylim=c(0, 100),
    censor.size=0.25,
    font.tickslab = c(12, "bold"),
    pval=TRUE,
    pval.size = 5,
    pval.coord = c(0,10),
    conf.int = FALSE) 
plot4
```
# FIGURE 2: Survival curves for medicaid enrollment variables 
```{r}
pdf("Figure 2.pdf", width=12, height=8.5, paper="USr")
plot1$plot  + plot2$plot + plot3$plot + plot4$plot + plot_layout(ncol=2, nrow=2, guides='keep') # & theme(legend.position='top')

dev.off()
```

# Get pair wise comparison p-values for survival curve differences
```{r}
pairwise_survdiff(Surv(fumo, can_dead) ~ mcont66.2, data = seer0to39 %>% filter(age_cat3 == "0 to 14"))
pairwise_survdiff(Surv(fumo, can_dead) ~ mcont66.2, data = seer0to39 %>% filter(age_cat3 == "15 to 39"))
```

# SUPPLEMENTARY TABLE 2: Get survival times overall and In_Medicaid_Flag
- 0 to 14 Overall
```{r}

SuppTable2a <-summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to39 %>% filter(age_cat3 == "0 to 14")), times=60) # estimates 60 month survival time

SuppTable2a <- as.data.frame(cbind(SuppTable2a$strata, SuppTable2a$time, round(SuppTable2a$surv*100, 2), round(SuppTable2a$lower*100, 2), round(SuppTable2a$upper*100,2))) # makes dataframe

SuppTable2a$Category <- "Overall"
SuppTable2a$Age_cat <- "0 to 14"

SuppTable2a <- setNames(SuppTable2a, c("Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Category", "Age group")) # gives column names

# Reorder  columns
SuppTable2a <- SuppTable2a[, c(5, 6,1:4)]
```

- 0 to 14 In_Medicaid_Flag
```{r}
# IN_MEDICAID_FLAG
SuppTable2b <-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to39 %>% filter(age_cat3 == "0 to 14")), times=60) # estimates 60 month survival time

SuppTable2b <- as.data.frame(cbind(SuppTable2b$strata, SuppTable2b$time, round(SuppTable2b$surv*100, 2), round(SuppTable2b$lower*100, 2), round(SuppTable2b$upper*100,2))) # makes dataframe

SuppTable2b <- SuppTable2b %>%
  mutate(V1 = case_when(V1 ==1 ~ "Not enrolled", V1 ==2 ~ "Enrolled"))

SuppTable2b$Age_cat <- "0 to 14"

SuppTable2b <- setNames(SuppTable2b, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Age group")) # gives column names

# Reorder  columns
SuppTable2b <- SuppTable2b[, c(1,6,2:5)]
```

- 15 to 39 Overall
```{r}

SuppTable2c<-summary(survfit(Surv(fumo, can_dead) ~ 1, data = seer0to39 %>% filter(age_cat3 == "15 to 39")), times=60) # estimates 60 month survival time

SuppTable2c <- as.data.frame(cbind(SuppTable2c$strata, SuppTable2c$time, round(SuppTable2c$surv*100, 2), round(SuppTable2c$lower*100, 2), round(SuppTable2c$upper*100,2))) # makes dataframe

SuppTable2c$Category <- "Overall"
SuppTable2c$Age_cat <- "15 to 39"

SuppTable2c <- setNames(SuppTable2c, c("Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Category", "Age group")) # gives column names

# Reorder  columns
SuppTable2c <- SuppTable2c[, c(5, 6,1:4)]
```

- 15 to 39 In_Medicaid_Flag
```{r}
SuppTable2d <-summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, data = seer0to39 %>% filter(age_cat3 == "15 to 39")), times=60) # estimates 60 month survival time

SuppTable2d <- as.data.frame(cbind(SuppTable2d$strata, SuppTable2d$time, round(SuppTable2d$surv*100, 2), round(SuppTable2d$lower*100, 2), round(SuppTable2d$upper*100,2))) # makes dataframe

SuppTable2d <- SuppTable2d %>%
  mutate(V1 = case_when(V1 ==1 ~ "Not enrolled",
                              V1 ==2 ~ "Enrolled"))

SuppTable2d$Age_cat <- "15 to 39"

SuppTable2d <- setNames(SuppTable2d, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Age group")) # gives column names

# Reorder  columns
SuppTable2d <- SuppTable2d[, c(1,6,2:5)]
```

#  SUPPLEMENTARY TABLE 2: Get survival times mcont66
- 0 to 14
```{r}
table(seer0to39$mcont66.2)
SuppTable2e <-summary(survfit(Surv(fumo, can_dead) ~ mcont66.2, data = seer0to39 %>% filter(age_cat3 == "0 to 14")), times=60) # estimates 60 month survival time

SuppTable2e <- as.data.frame(cbind(SuppTable2e$strata, SuppTable2e$time, round(SuppTable2e$surv*100, 2), round(SuppTable2e$lower*100, 2), round(SuppTable2e$upper*100,2))) # makes dataframe

SuppTable2e <- SuppTable2e %>%
  mutate(V1 = case_when(V1 == 1 ~ "Not enrolled",
                        V1 == 2 ~ "Continuous",
                        V1 == 3 ~ "Discontinuous",
                        V1 == 4 ~ "Other",
                        V1 == 5 ~ "At diagnosis"))

SuppTable2e$Age_cat <- "0 to 14"

SuppTable2e <- setNames(SuppTable2e, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Age group")) # gives column names

# Reorder  columns
SuppTable2e <- SuppTable2e[, c(1,6,2:5)]

# delete not enrolled row
SuppTable2e <- SuppTable2e %>%
  filter(Category != "Not enrolled")
```

- 15-39
```{r}
SuppTable2f <-summary(survfit(Surv(fumo, can_dead) ~ mcont66.2, data = seer0to39 %>% filter(age_cat3 == "15 to 39")), times=60) # estimates 60 month survival time

# 20 to 39
SuppTable2f <- as.data.frame(cbind(SuppTable2f$strata, SuppTable2f$time, round(SuppTable2f$surv*100, 2), round(SuppTable2f$lower*100, 2), round(SuppTable2f$upper*100,2))) # makes dataframe

SuppTable2f <- SuppTable2f %>%
  mutate(V1 = case_when(V1 == 1 ~ "Not enrolled",
                        V1 == 2 ~ "Continuous",
                        V1 == 3 ~ "Discontinuous",
                        V1 == 4 ~ "Other",
                        V1 == 5 ~ "At diagnosis"
                        ))

SuppTable2f$Age_cat <- "15 to 39"

SuppTable2f <- setNames(SuppTable2f, c("Category", "Time (months)", "Survival Percentage", "Lower CI", "Upper CI", "Age group")) # gives column names

# Reorder  columns
SuppTable2f <- SuppTable2f[, c(1,6,2:5)]

# delete not enrolled row
SuppTable2f <- SuppTable2f %>%
  filter(Category != "Not enrolled")
```

# SUPPLEMENTARY TABLE 2
```{r}
SuppTable2 <- rbind(SuppTable2a, SuppTable2b, SuppTable2c, SuppTable2d, SuppTable2e, SuppTable2f)
```

# 60 month survival by CNS tumor type
```{r}
# Initialize an empty data frame with appropriate column names
columns <- c("Medicaid_Category", "Survival Percentage", "LowerCI", "UpperCI", "Brain tumor subtype")
results <- data.frame(matrix(ncol = length(columns), nrow = 0))
colnames(results) <- columns

ct <- unique(seer0to39$brain2)

for (i in ct) {
  df <- seer0to39 %>% filter(brain2 == i, age_cat3 == "0 to 14")
  
  # Calculate survival at 60 months
  x <- summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times = 60)
  
  # Create a new row for the result
  new_row <- data.frame(
    Medicaid_Category = x$strata,
    Survival_Percentage = round(x$surv, 3) * 100,
    LowerCI = round(x$lower, 3) * 100,
    UpperCI = round(x$upper, 3) * 100,
    Brain_tumor_subtype = i
  )
  
  # Append the new row to the results data frame
  results <- rbind(results, new_row)
}

# Optionally, you can sort the results by a specific column
results <- results[order(results$Brain_tumor_subtype), ]

# Add labels to Medicaid categories
SuppTable3a <- results %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == "In_Medicaid_Flag=Not Enrolled" ~ "Not enrolled",
                              Category == "In_Medicaid_Flag=Enrolled" ~ "Enrolled")) %>%
  mutate(Age_cat = "0 to 14")


# In_Medicaid_Flag 15 to 39

# Initialize an empty data frame with appropriate column names
columns <- c("Medicaid_Category", "Survival Percentage", "LowerCI", "UpperCI", "Brain tumor subtype")
results <- data.frame(matrix(ncol = length(columns), nrow = 0))
colnames(results) <- columns

ct <- unique(seer0to39$brain2)

for (i in ct) {
  df <- seer0to39 %>% filter(brain2 == i, age_cat3 == "15 to 39")
  
  # Calculate survival at 60 months
  x <- summary(survfit(Surv(fumo, can_dead) ~ In_Medicaid_Flag, df), times = 60)
  
  # Create a new row for the result
  new_row <- data.frame(
    Medicaid_Category = x$strata,
    Survival_Percentage = round(x$surv, 3) * 100,
    LowerCI = round(x$lower, 3) * 100,
    UpperCI = round(x$upper, 3) * 100,
    Brain_tumor_subtype = i
  )
  
  # Append the new row to the results data frame
  results <- rbind(results, new_row)
}

# Optionally, you can sort the results by a specific column
results <- results[order(results$Brain_tumor_subtype), ]

# Add labels to Medicaid categories
SuppTable3b <- results %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == "In_Medicaid_Flag=Not Enrolled" ~ "Not enrolled",
                              Category == "In_Medicaid_Flag=Enrolled" ~ "Enrolled")) %>%
  mutate(Age_cat = "15 to 39")
```

# 60 month survival by CNS tumor type mcont66.2
```{r}
# Initialize an empty data frame with appropriate column names
columns <- c("Medicaid_Category", "Survival Percentage", "LowerCI", "UpperCI", "Brain tumor subtype")
results <- data.frame(matrix(ncol = length(columns), nrow = 0))
colnames(results) <- columns

ct <- unique(seer0to39$brain2)

for (i in ct) {
  df <- seer0to39 %>% filter(brain2 == i, age_cat3 == "0 to 14")
  
  # Calculate survival at 60 months
  x <- summary(survfit(Surv(fumo, can_dead) ~ mcont66.2, df), times = 60)
  
  # Create a new row for the result
  new_row <- data.frame(
    Medicaid_Category = x$strata,
    Survival_Percentage = round(x$surv, 3) * 100,
    LowerCI = round(x$lower, 3) * 100,
    UpperCI = round(x$upper, 3) * 100,
    Brain_tumor_subtype = i
  )
  
  # Append the new row to the results data frame
  results <- rbind(results, new_row)
}

# Optionally, you can sort the results by a specific column
results <- results[order(results$Brain_tumor_subtype), ]

# Add labels to Medicaid categories
SuppTable3c <- results %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == "mcont66.2=Not enrolled" ~ "Not enrolled",
                              Category == "mcont66.2=Continuous" ~ "Continuous",
                              Category == "mcont66.2=Discontinuous" ~ "Discontinuous",
                              Category == "mcont66.2=Other" ~ "Other",
                              Category == "mcont66.2=At diagnosis" ~ "At diagnosis")) %>%
  mutate(Age_cat = "0 to 14")


# In_Medicaid_Flag 15 to 39

# Initialize an empty data frame with appropriate column names
columns <- c("Medicaid_Category", "Survival Percentage", "LowerCI", "UpperCI", "Brain tumor subtype")
results <- data.frame(matrix(ncol = length(columns), nrow = 0))
colnames(results) <- columns

for (i in ct) {
  df <- seer0to39 %>% filter(brain2 == i, age_cat3 == "15 to 39")
  
  # Calculate survival at 60 months
  x <- summary(survfit(Surv(fumo, can_dead) ~ mcont66.2, df), times = 60)
  
  # Create a new row for the result
  new_row <- data.frame(
    Medicaid_Category = x$strata,
    Survival_Percentage = round(x$surv, 3) * 100,
    LowerCI = round(x$lower, 3) * 100,
    UpperCI = round(x$upper, 3) * 100,
    Brain_tumor_subtype = i
  )
  
  # Append the new row to the results data frame
  results <- rbind(results, new_row)
}

# Optionally, you can sort the results by a specific column
results <- results[order(results$Brain_tumor_subtype), ]

# Add labels to Medicaid categories
SuppTable3d <- results %>%
  rename(Category = Medicaid_Category) %>%
  mutate(Category = case_when(Category == "mcont66.2=Not enrolled" ~ "Not enrolled",
                              Category == "mcont66.2=Continuous" ~ "Continuous",
                              Category == "mcont66.2=Discontinuous" ~ "Discontinuous",
                              Category == "mcont66.2=Other" ~ "Other",
                              Category == "mcont66.2=At diagnosis" ~ "At diagnosis")) %>%
  mutate(Age_cat = "15 to 39")
```

# SUPPLEMENTARY TABLE 3: 60 month survival by CNS tumor type 
```{r}
SuppTable3 <- rbind(SuppTable3a, SuppTable3b, SuppTable3c, SuppTable3d)

table(SuppTable3$Brain_tumor_subtype)
# Abbreviate brain tumor subtype

SuppTable3 <- SuppTable3 %>%
  mutate(abbrev = case_when(Brain_tumor_subtype == "3.1 Astrocytoma" ~ 0,
                            Brain_tumor_subtype == "3.2 Other glioma" ~ 1,
                            Brain_tumor_subtype == "3.3 Ependymoma" ~ 2,
                            Brain_tumor_subtype == "3.4 Medulloblastoma and other PNET" ~ 3,
                            Brain_tumor_subtype == "Other specified (3.5) and unspecified (3.6) intracranial and instraspinal neoplasms" ~ 4))

SuppTable3$abbrev <- factor(SuppTable3$abbrev, levels = c(0:4), labels = c("ASTRO", "OGLI", "EPEN", "MB/PNET", "OSPEC/UNSPEC"))
```

# FIGURE 3: Plot 60 month survival by tumor type and age group
```{r}
# 0 to 19
Plot1 <- SuppTable3 %>%
  filter(Age_cat == "0 to 14" & Category %in% c("Enrolled", "Not enrolled")) %>%
  ggplot() +
  geom_point(aes(y = Survival_Percentage, x = abbrev, group = Category, color = Category), size=2, position = position_dodge(width=0.9)) + 
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y = Survival_Percentage, x = abbrev, ymax = LowerCI, ymin = UpperCI, group = Category, color = Category), height =0.5, width = 0.4, position = position_dodge(width=0.9)) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=12, face = "bold", color = "black"),
    axis.text.y=element_text(size=12, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, x= "Brain tumor subtype", y = "60-month survival (%)", title = "B. 0 to 14 years") +
  ylim(0, 100) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold")) 

Plot1

# 20 to 39
Plot2 <- SuppTable3 %>%
  filter(Age_cat == "15 to 39" & Category %in% c("Enrolled", "Not enrolled")) %>%
  ggplot() +
  geom_point(aes(y = Survival_Percentage, x = abbrev, group = Category, color = Category), size=2, position = position_dodge(width=0.9)) + 
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y = Survival_Percentage, x = abbrev, ymax = LowerCI, ymin = UpperCI, group = Category, color = Category), height =0.5, width = 0.4, position = position_dodge(width=0.9)) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=12, face = "bold", color = "black"),
    axis.text.y=element_text(size=12, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, x= "Brain tumor subtype", y = "60-month survival (%)", title = "B. 15 to 39 years") +
  ylim(0, 100) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold")) 

Plot2

# 0 to 19
Plot3 <- SuppTable3 %>%
  filter(Age_cat == "0 to 14" & !Category %in% c("Enrolled")) %>%
  ggplot() +
  geom_point(aes(y = Survival_Percentage, x = abbrev, group = Category, color = Category), size=2, position = position_dodge(width=0.9)) + 
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y = Survival_Percentage, x = abbrev, ymax = LowerCI, ymin = UpperCI, group = Category, color = Category), height =0.5, width = 0.4, position = position_dodge(width=0.9)) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=12, face = "bold", color = "black"),
    axis.text.y=element_text(size=12, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, x= "Brain tumor subtype", y = "60-month survival (%)", title = "C. 0 to 14 years") +
  ylim(0, 100) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold"))  +
  guides(color = guide_legend(nrow = 3, byrow = TRUE))

Plot3

# 20 to 39
Plot4 <- SuppTable3 %>%
  filter(Age_cat == "15 to 39" & !Category %in% c("Enrolled")) %>%
  ggplot() +
  geom_point(aes(y = Survival_Percentage, x = abbrev, group = Category, color = Category), size=2, position = position_dodge(width=0.9)) + 
  geom_vline(xintercept = c(1.5, 2.5, 3.5, 4.5, 5.5), linetype=2) +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y = Survival_Percentage, x = abbrev, ymax = LowerCI, ymin = UpperCI, group = Category, color = Category), height =0.5, width = 0.4, position = position_dodge(width=0.9)) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=12, face = "bold", color = "black"),
    axis.text.y=element_text(size=12, face = "bold", color = "black"),
    axis.title.x=element_text(size=12, face = "bold", color = "black"),
    axis.title.y=element_text(size=12, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL, x= "Brain tumor subtype", y = "60-month survival (%)", title = "D. 15 to 39 years") +
  ylim(0, 100) +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 9, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=12, face="bold")) +
   guides(color = guide_legend(nrow = 3, byrow = TRUE))

Plot4
```
# FIGURE 3: 60 months survival plots
```{r}
pdf("Figure 3.pdf", width=11, height=8.5, paper="USr")
Plot1 + Plot2 + Plot3 + Plot4  + plot_layout(ncol=2, nrow=2) 

dev.off()
```

```{r}
table(seer0to39$mcont66.2)
```

# TABLE 3: Cox proportional hazards models for In_Medicaid_Flag and mcont66.2
```{r}
summaryn <- seer0to39 %>%
  group_by(age_cat2, brain2, mcont66.2)%>%
  summarize(n = n()) 

# adjusted models 0 to 14
res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 
res.cox2
summary(res.cox2)


Table3a<-as.data.frame(exp(cbind(HR = coef(res.cox2), confint(res.cox2))))
   Table3a <- Table3a[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race/ethnicity, age, and Yost SES indicator") %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled")) 
   
Table3a$`Age category` <- "0 to 14"


# adjusted models 15 to 39
res.cox3 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 
res.cox3
summary(res.cox3)

Table3b<-as.data.frame(exp(cbind(HR = coef(res.cox3), confint(res.cox3))))
   Table3b <- Table3b[c(1),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race/ethnicity, age, and Yost SES indicator") %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "In_Medicaid_FlagEnrolled" ~ "Enrolled")) 
   
Table3b$`Age category` <- "15 to 39"

# adjusted models 0 to 14
res.cox4 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 
summary(res.cox4)

table(seer0to39$mcont66)

Table3c<-as.data.frame(exp(cbind(HR = coef(res.cox4), confint(res.cox4))))

Table3c <- Table3c[c(1,2,3,4),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race/ethnicity, age, and Yost SES indicator") %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "mcont66.2Discontinuous" ~ "Discontinuous",
                                Category == "mcont66.2Other" ~ "Other",
                                Category == "mcont66.2Continuous" ~ "Continuous",
                                Category == "mcont66.2At diagnosis" ~ "At diagnosis"))

Table3c$`Age category` <- "0 to 14"

# adjusted models 15 to 39
res.cox5 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 
summary(res.cox5)


Table3d<-as.data.frame(exp(cbind(HR = coef(res.cox5), confint(res.cox5))))

Table3d <- Table3d[c(1,2,3,4),] %>%
    rownames_to_column() %>%
    mutate(model="Adjusted for race/ethnicity, age, and Yost SES indicator") %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "mcont66.2Discontinuous" ~ "Discontinuous",
                                Category == "mcont66.2Other" ~ "Other",
                                Category == "mcont66.2Continuous" ~ "Continuous",
                                Category == "mcont66.2At diagnosis" ~ "At diagnosis"))

Table3d$`Age category` <- "15 to 39"


Table3 <-rbind(Table3a, Table3b, Table3c, Table3d)
```


# SUPPLEMENTARY TABLE 4: Cox proportional hazards models for In_Medicaid_Flag and mcont66.2 by the tumor subtype
```{r}
age_cat <- c("0 to 14", "15 to 39")

# Initialize an empty data frame to store results
results <- data.frame()

# Nested loops for ct, age_cat2, and variables
for (i in ct) {
  for (j in age_cat) {
    for (variable in c("In_Medicaid_Flag", "mcont66.2")) {
      df <- seer0to39 %>% filter(brain2 == i, age_cat3 == j)
      
      # adjusted models 0 to 19
      formula <- as.formula(paste("Surv(fumo, can_dead) ~", variable, "+ race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state)"))
      res.cox2 <- coxph(formula, data = df) 
      print(paste("Results for ct =", i, ", age_cat2 =", j, ", and variable =", variable))
      print(summary(res.cox2))

      df2 <- as.data.frame(tidy(res.cox2, exponentiate = TRUE, conf.int = TRUE))
      df2$btype <- i
      df2$age_cat <- j
      df2$variable <- variable

      # Append the new row to the results data frame
      results <- rbind(results, df2 %>% filter(term %in% c("In_Medicaid_FlagEnrolled", "mcont66.2Continuous", "mcont66.2Discontinuous", "mcont66.2Other", "mcont66.2At diagnosis")) %>% select(term, estimate, conf.low, conf.high, btype, age_cat, variable))
    }
  }
}

SuppTable4 <- results
SuppTable4$abbrev <- factor(SuppTable4$btype, levels = c("3.1 Astrocytoma", "3.2 Other glioma", "3.3 Ependymoma", "3.4 Medulloblastoma and other PNET", "Other specified (3.5) and unspecified (3.6) intracranial and instraspinal neoplasms"), labels = c("ASTRO", "OGLI", "EPEN", "MB/PNET", "OTHER"))

# Strip off "mcont66.2" from the variable values
SuppTable4$term <- sub("^mcont66.2", "", SuppTable4$term)

# Strip off "In_Medicaid_Flag" from the variable values
SuppTable4$term <- sub("^In_Medicaid_Flag", "", SuppTable4$term)

# remove other
SuppTable4 <- SuppTable4 %>%
  filter(abbrev != "OTHER")

table(SuppTable4$btype)
```

```{r}
# Create a function to generate plots
plot_survival_results <- function(data, variable, age_group, title) {
  df_subset <<- data %>% filter(variable == variable, age_cat == age_group)

  # Check if the subset is not empty
  if (nrow(df_subset) > 0) {
    p <- ggplot(df_subset, aes(x = estimate, y = abbrev, color = term, group=term)) +
      geom_point(position = position_dodge(width = 0.8), size = 3) +
      geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.5, position = position_dodge(width = 0.8)) +
      geom_vline(xintercept =1, linetype=2) +
      labs(title = title,
           x = "HR", y = "Brain Tumor Subtype") +
      theme_classic() +
      theme(legend.position = "bottom",
            axis.title.x=element_text(size=12, face = "bold", color = "black"),
            axis.title.y=element_text(size=12, face = "bold", color = "black"),
            plot.title = element_text(face="bold")) +
  theme(legend.title=element_blank())
   guides(color = guide_legend(nrow = 3, byrow = TRUE))

    print(p)
  }
}

A <- plot_survival_results(SuppTable4 %>% filter(variable == "In_Medicaid_Flag"), variable, "0 to 14", title = "A. 0 to 14")
B <- plot_survival_results(SuppTable4 %>% filter(variable == "In_Medicaid_Flag"), variable, "15 to 39", title = "B. 15 to 39")
C <- plot_survival_results(SuppTable4 %>% filter(variable == "mcont66.2"), variable, "0 to 14", title = "C. 0 to 14")
D <- plot_survival_results(SuppTable4 %>% filter(variable == "mcont66.2"), variable, "15 to 39", title = "C. 15 to 39")
```

# SUPPLEMENTARY FIGURE 1
```{r}
pdf("SUPPLEMENTARY FIGURE 1.pdf", width=11, height=8.5, paper="USr")
A + B + C + D  + plot_layout(ncol=2, nrow=2) 

dev.off()
```


# Test for effect modification by race, sex
- 0 to 14
```{r}
# race/ethnicity 0 to 19 In_Medicaid_Flag
race <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 

raceem<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + race_ethnicity*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3== "0 to 14"))
lrtest(race, raceem)

# mcont66.2
race2 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 

raceem2<-coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + race_ethnicity*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14"))

lrtest(race2, raceem2)

# No significant effect modification by race/ethnicity for 0 to 14 year olds
```

- 15 to 39
```{r}
race <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile, data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 

raceem<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + race_ethnicity*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 
lrtest(race, raceem)

# mcont66.2
race2 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile, data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 

raceem2 <-coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + race_ethnicity*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 
lrtest(race2, raceem2)

# EM by race/ethnicity category present for 15 to 39 year olds
```

- sex
- 0 to 14
```{r}
sex <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 

sexem<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + sex*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 
lrtest(sex, sexem)

sex2 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 

sexem2<-coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + sex*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "0 to 14")) 
lrtest(sex2, sexem2)

# No effect modification by sex for 0 to 14 year olds
```

- 15 to 39
```{r}
sex <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 

sexem<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + sex*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 
lrtest(sex, sexem)

sex2 <- coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 

sexem2<-coxph(Surv(fumo, can_dead) ~ mcont66.2 + race_ethnicity + age + Yost_ACS_2006_2010_quintile + sex + sex*In_Medicaid_Flag + cluster(state), data = seer0to39 %>% filter(age_cat3 == "15 to 39")) 
lrtest(sex2, sexem2)

# EM by sex category present for In_Medicaid_Flag for 15-39 year olds
```


# SUPPLEMENTARY TABLE 5a: Cox proportional hazards models by sex and race/ethnicity for 20 to 39 year olds
```{r}
run_cox_model <- function(Sex, var, ac) {
  x <<- seer0to39 %>%
    filter(sex == Sex, age_cat3 == ac)
  
  formula_str <- paste("Surv(fumo, can_dead) ~", var, "+ age + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state)")
  model <- coxph(as.formula(formula_str), data = x)
  
  a <- as.data.frame(exp(cbind(HR = coef(model), confint(model))))
  
  if (nrow(a) == 1) {
    a <- a[1,] %>%
      rownames_to_column() %>%
      mutate(model = paste("Adjusted for age, sex,  and Yost SES indicator", sep = " "), Group = Sex, agegroup = ac)
  } else {
    a <- a[1:4,] %>%
      rownames_to_column() %>%
      mutate(model = paste("Adjusted for age, sex and Yost SES index", sep = " "), Group = Sex, agegroup = ac)
  }
  
  # Keep only the results for In_Medicaid_Flag and mcont66.2
  a <- a %>% filter(rowname %in% c("In_Medicaid_FlagEnrolled", "mcont66.2Discontinuous", "mcont66.2Continuous", "mcont66.2Other", "mcont66.2At diagnosis")) %>%
    mutate(across(c(2:4), ~ round(., 2)))
  
  return(a)
}

categories <- c("Male", "Female")
variables <- c("In_Medicaid_Flag", "mcont66.2")
age_groups <- c("0 to 14", "15 to 39")

SuppTable5a <- data.frame()

for (Sex in categories) {
  for (var in variables) {
    for (ac in age_groups) {
      result <- run_cox_model(Sex, var, ac)
      result$variable <- var
      result$variable2 <- "sex"
      SuppTable5a <- bind_rows(SuppTable5a, result)
    }
  }
}

options(scipen = 999)

SuppTable5a$rowname <- sub("^mcont66.2", "", SuppTable5a$rowname)
SuppTable5a$rowname <- sub("^In_Medicaid_Flag", "", SuppTable5a$rowname)
SuppTable5a <- SuppTable5a %>%
  rename("Category" = "rowname") %>%
  filter(agegroup == "15 to 39", variable == "In_Medicaid_Flag") # 0 to 14 was not significant so remove
```

# SUPPLEMENTARY TABLE 5b: Cox proportional hazards models by race/ethnicity for 15 to 39 year olds
```{r}
run_cox_model <- function(re, var, ac) {
  x <- seer0to39 %>%
    filter(race_ethnicity == re, age_cat3 == ac)
  
  formula_str <- paste("Surv(fumo, can_dead) ~", var, "+ age + Yost_ACS_2006_2010_quintile + cluster(state)")
  model <- coxph(as.formula(formula_str), data = x)
  
  a <- as.data.frame(exp(cbind(HR = coef(model), confint(model))))
  
  if (nrow(a) == 1) {
    a <- a[1,] %>%
      rownames_to_column() %>%
      mutate(model = paste("Adjusted for age and Yost SES indicator", sep = " "), Group = re, agegroup = ac)
  } else {
    a <- a[1:4,] %>%
      rownames_to_column() %>%
      mutate(model = paste("Adjusted for age and Yost SES index", sep = " "), Group = re, agegroup = ac)
  }
  
  # Keep only the results for In_Medicaid_Flag and mcont66.2
  a <- a %>% filter(rowname %in% c("In_Medicaid_FlagEnrolled", "mcont66.2Discontinuous", "mcont66.2Continuous", "mcont66.2Other", "mcont66.2At diagnosis")) %>%
    mutate(across(c(2:4), ~ round(., 2)))
  
  return(a)
}

categories <- c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic API", "Non-Hispanic Other", "Hispanic")
variables <- c("In_Medicaid_Flag", "mcont66.2")
age_groups <- c("0 to 14", "15 to 39")

SuppTable5b <- data.frame()

for (re in categories) {
  for (var in variables) {
    for (ac in age_groups) {
      result <- run_cox_model(re, var, ac)
      result$variable <- var
      result$variable2 <- "race_ethnicity"
      SuppTable5b <- bind_rows(SuppTable5b, result)
    }
  }
}

options(scipen = 999)

SuppTable5b$rowname <- sub("^mcont66.2", "", SuppTable5b$rowname)
SuppTable5b$rowname <- sub("^In_Medicaid_Flag", "", SuppTable5b$rowname)
SuppTable5b <- SuppTable5b %>%
  rename("Category" = "rowname") %>%
  filter(agegroup == "15 to 39") # 0 to 14 was not signficant so remove
```
# SUPPLEMENTARY TABLE 5
```{r}
SuppTable5 <- rbind(SuppTable5a, SuppTable5b)
```

# SUPPLEMENTARY FIGURE 2 
```{r}
# Create a function to generate plots
plot_survival_results <- function(data, Variable, age_group, title, Variable2, ytitle) {
  df_subset <<- data %>% filter(variable == Variable, agegroup == age_group, variable2 == Variable2)

  # Check if the subset is not empty
  if (nrow(df_subset) > 0) {
    p <- ggplot(df_subset, aes(x = HR, y = Group, color = Category, group=Category)) +
      geom_point(position = position_dodge(width = 0.8), size = 2) +
      geom_errorbarh(aes(xmin =`2.5 %`, xmax = `97.5 %`), height = 0.5, position = position_dodge(width = 0.8)) +
      geom_vline(xintercept =1, linetype=2) +
      labs(title = title,
           x = "HR", y = ytitle) +
      theme_classic() +
      theme(legend.position = "bottom",
            axis.title.x=element_text(size=12, face = "bold", color = "black"),
            axis.title.y=element_text(size=12, face = "bold", color = "black"),
            plot.title = element_text(face="bold")) +
  theme(legend.title=element_blank()) 

    print(p)
  }
}

A <- plot_survival_results(SuppTable5, "In_Medicaid_Flag", "15 to 39",  title = "A. 15 to 39", "sex", "Sex")
B <- plot_survival_results(SuppTable5, "mcont66.2", "15 to 39",  title = "B. 15 to 39", "sex", "Sex")
C <- plot_survival_results(SuppTable5, "In_Medicaid_Flag", "15 to 39",  title = "C. 15 to 39", "race_ethnicity", "Race/ethnicity category")
D <- plot_survival_results(SuppTable5, "mcont66.2", "15 to 39",  title = "D. 15 to 39", "race_ethnicity", "Race/ethnicity category")


pdf("SUPPLEMENTARY FIGURE 2.pdf", width=11, height=8.5, paper="USr")
A + B + C + D  + plot_layout(ncol=2, nrow=2) 

dev.off()

```

# Cox PH assumption testing by age group (0-14) NOT VIOLATED
```{r}
# In_Medicaid_Flag
# adjusted models 0 to 19
seer0to14 <- seer0to39 %>% filter(age_cat3 == "0 to 14")
    
model0to14<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag  + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to14) # Model adjusted for race, age category, and a measure of poverty

Medflag0to14 <- cox.zph(model0to14, terms = FALSE)
Medflag0to14

plot(Medflag0to14[1], col=c("black", "red"), main="Medicaid_flag log log plot", fun="cloglog", xlab="Time", ylab="log(-log(S(t))") #  ph assumption violated

plot(Medflag0to14, resid = TRUE)

ggcoxdiagnostics(model0to14, type = "schoenfeld") 
```

# Cox PH assumption testing by age group (15-39) NOT VIOLATED
```{r}
# In_Medicaid_Flag
# adjusted models 20 to 39
seer15to39 <- seer0to39 %>% filter(age_cat3 == "15 to 39")
    
model15to39<-coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag  + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = seer15to39) # Model adjusted for race, age category, and a measure of poverty

Medflag15to39 <- cox.zph(model15to39, terms = FALSE)
Medflag15to39

plot(Medflag15to39[1], col=c("black", "red"), main="Medicaid_flag log log plot", fun="cloglog", xlab="Time", ylab="log(-log(S(t))") #  ph assumption violated

plot(Medflag15to39, resid = TRUE)

ggcoxdiagnostics(model15to39, type = "schoenfeld") 
```

# Run models in 12 month increments because 15 to 39 was close to violation
```{r}
# Split a survival data set at specified times to form a counting process format
# For those with survival time of 0, add 0.5 mo because otherwise splitting doesn't work and it is more accurate
seer0to39 <- seer0to39 %>%
  mutate(fumo_half = if_else(fumo == 0, 0.5, fumo))

seer0to39$patient_id <- as.numeric(rownames(seer0to39))

seer0to39.cp.format <- survSplit(data  = seer0to39,
                            cut   = c(12, 24, 36),  # cut at time 12,24,36
                            end   = "fumo_half",    # original survival time
                            event = "can_dead",     # event indicator
                            start = "start", # will be created. zero by default
                            episode ="timegroup")    

# Recoding
seer0to39.cp.format <- within(seer0to39.cp.format, {
    # Create new survival object
    SurvObj <- Surv(start, fumo_half, can_dead) # creates column labeled col
    # Create interval indicator
    interval <- factor(start, levels = c(0, 12, 24, 36), labels = c("First","Second", "Third", "Fourth"))
})

# Reordering
seer0to39.cp.format <- seer0to39.cp.format[with(seer0to39.cp.format, order(patient_id, fumo_half)),]  # sort by patient_id and then by fumo
summary(seer0to39$fumo_half)
```

# for loop for models by intervals with In_Medicaid_Flag, mcont66, for 0 to 19 and 20 to 39 overall and by cancer type
```{r}
seer0to39.cp.format <- seer0to39.cp.format %>%
  mutate(age_cat3 = case_when(age_cat3 == "0 to 14" ~ 0,
                              age_cat3 == "15 to 39" ~ 1)) %>%
  # create can_type age group categories
  mutate(age_can = case_when(age_cat3 == 0 & brain == "3.1 Astrocytoma" ~ 0,
                             age_cat3 == 0 & brain == "3.2 Other glioma"  ~ 1,
                             age_cat3 == 1 & brain == "3.1 Astrocytoma" ~ 2,
                             age_cat3 == 1 & brain == "3.2 Other glioma"  ~ 3))

seer0to39.cp.format$age_cat3 <- factor(seer0to39.cp.format$age_cat3, levels = c(0:1), labels = c("0 to 14", "15 to 39"))

seer0to39.cp.format$age_can<- factor(seer0to39.cp.format$age_can, levels = c(0:3), labels = c("0-14, 3.1 Astrocytoma", "0-14, 3.2 Other glioma", "15-39, 3.1 Astrocytoma", "15-39, 3.2 Other glioma"))

table(seer0to39.cp.format$age_cat3)
table(seer0to39.cp.format$age_can)

age <- unique(seer0to39.cp.format$age_cat3)


# testing here
m <-coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + age + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), seer0to39.cp.format[which(seer0to39.cp.format$age_cat3 == "15 to 39"),])

m

X<- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0), 1)

t <- glht(m, linfct = X) 
summary(t) # here are the results

CI <- confint(t)
CI <- as.data.frame(exp(CI$confint))
# end testing

# In_Medicaid_Flag overall
results <- data.frame(matrix(ncol=6, nrow=0)) # create empty table for results that will be populated by for loop
columns<-c("estimate", "lwr", "upr", "age", "interval", "cancer_type") # label the columns with names
colnames(results) <- columns

W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
X <- matrix(c(1,0,0,0,0,0,0,0,0,0,1,0,0), 1)
Y <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,1,0), 1)
Z <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1), 1)

for (i in age) {
  df <- seer0to39.cp.format %>% filter(age_cat3 == i)
  x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + age + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
  t0<- glht(x, linfct = W) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))
  df1$age <- i
  df1$interval = "0-12" # enrolled first interval
  df1$cancer_type = "Overall"
  
  t12<- glht(x, linfct = X) 
  df2<-confint(t12)
  df2 <- as.data.frame(exp(df2$confint))
  df2$age <- i
  df2$interval = ">12-24" # enrolled second interval
  df2$cancer_type = "Overall"
  
  t24<- glht(x, linfct = Y) 
  df3<-confint(t24)
  df3 <- as.data.frame(exp(df3$confint))
  df3$age <- i
  df3$interval = ">24-36" # enrolled third interval
  df3$cancer_type = "Overall"
  
  t36 <- glht(x, linfct = Z) 
  df4 <-confint(t36)
  df4 <- as.data.frame(exp(df4$confint))
  df4$age <- i
  df4$interval = "36" # enrolled fourth interval
  df4$cancer_type = "Overall"
  results <-rbind(results, df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
}

results$category = "Enrolled"
PHresults.enrolled <- results
```

# By age and cancer type 
```{r}
# In_Medicaid_Flag overall
results <- data.frame(matrix(ncol=5, nrow=0)) # create empty table for results that will be populated by for loop
columns<-c("estimate", "lwr", "upr",  "interval", "age_can") # label the columns with names
colnames(results) <- columns

W <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0), 1) 
X <- matrix(c(1,0,0,0,0,0,0,0,0,0,1,0,0), 1)
Y <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,1,0), 1)
Z <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1), 1)
table(seer0to39.cp.format$age_can)

df_filtered <- seer0to39.cp.format %>% 
  filter(age_can == "0-14, 3.1 Astrocytoma"|age_can == "15-39, 3.1 Astrocytoma"|age_can == "0-14, 3.2 Other glioma"|age_can == "15-39, 3.2 Other glioma")

df_filtered$age_can <- droplevels(df_filtered$age_can)


table(df_filtered$age_can)
age_can <- unique(df_filtered$age_can)
age_can

for (i in age_can) {
  df <- df_filtered %>% filter(age_can == i) 
  
  x <- coxph(SurvObj ~ In_Medicaid_Flag + In_Medicaid_Flag*strata(interval) + age + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
  t0<- glht(x, linfct = W) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))
  df1$interval = "0-12" # enrolled first interval
  df1$age_can = i
  
  t12<- glht(x, linfct = X) 
  df2<-confint(t12)
  df2 <- as.data.frame(exp(df2$confint))
  df2$interval = "12-24" # enrolled second interval
  df2$age_can = i
  
  t24<- glht(x, linfct = Y) 
  df3<-confint(t24)
  df3 <- as.data.frame(exp(df3$confint))
  df3$interval = "24-36" # enrolled third interval
  df3$age_can = i
  
  t36 <- glht(x, linfct = Z) 
  df4 <-confint(t36)
  df4 <- as.data.frame(exp(df4$confint))
  df4$interval = "36+" # enrolled fourth interval
  df4$age_can = i
  results <-rbind(results, df1, df2, df3, df4)  # this puts the results from each loop round in the results data file
}

results$category = "Enrolled"
PHresults.enrolled2 <- results
```

# SUPPLEMENTARY TABLE 6: PH Results
- combine PHresults.enrolled and PHresults.enrolled2
```{r}
# make age_can variable for PHresults.enrolled

age_can <- rep(c("15 to 39, Overall", "0 to 14, Overall"), each = 4)
age_can

PHresults.enrolled <- cbind(PHresults.enrolled, age_can)

PHresults.enrolled <- PHresults.enrolled %>%
  select(-age, -cancer_type)

PHresults.enrolled <- PHresults.enrolled %>%
  select(Estimate, lwr, upr, interval, age_can, category)

SuppTable6 <- rbind(PHresults.enrolled, PHresults.enrolled2)
```

# Graph hazard ratios and confidence intervals across time for enrolled variable
```{r}
PHresults.enrolled$interval <-  factor(PHresults.enrolled$interval, levels = c("0-12", ">12-24", ">24-36", "36"))
# graph HRs and 95% CIs

p<- PHresults.enrolled %>%
  ggplot() +
  geom_point(aes(y=Estimate, x= interval, group=age_can, color = age_can), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr, group=age_can, color = age_can), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=16, face = "bold", color = "black"),
    axis.text.y=element_text(size=16, face = "bold", color = "black"),
    axis.title.x=element_text(size=20, face = "bold", color = "black"),
    axis.title.y=element_text(size=20, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Time interval") +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 15, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=15, face="bold")) 

p
```
# mcont66.2
```{r}
# testing here
m <-coxph(SurvObj ~ mcont66.2 + mcont66.2*strata(interval) + age + race_ethnicity +  Yost_ACS_2006_2010_quintile + cluster(state), data = seer0to39.cp.format[which(seer0to39.cp.format$age_cat3 == '0 to 14'),]) 
m

X <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), 1)

t <- glht(m, linfct = X) 
summary(t) # here are the results

CI<-confint(t)
CI <- as.data.frame(exp(CI$confint))
# end testing

# mcont66.2
results <- data.frame(matrix(ncol=5, nrow=0)) # create empty table for results that will be populated by for loop
columns<-c("estimate", "lwr", "upr", "age", "interval") # label the columns with names
colnames(results) <- columns

A <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,0,0,0,0,0,0), 1) # Continuous 1st interval
B <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,0,0,0,0,0,0), 1) # Discontinuous 1st interval
C <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,0,0,0,0,0,0), 1) # Other 1st interval
C.2 <- matrix(c(0,0,0,1,0,0,0,0,0,0,0,0,0, 0,0,0,0,0,0,0,0,0,0,0,0), 1) # At diagnosis 1st interval

D <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,  1,0,0,0,0,0,0,0,0,0,0,0), 1) # Continuous 2nd interval
E <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,  0,1,0,0,0,0,0,0,0,0,0,0), 1) # Discontinuous 2nd interval
F <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,  0,0,1,0,0,0,0,0,0,0,0,0), 1) # Other 2nd interval
F.2 <- matrix(c(0,0,0,1,0,0,0,0,0,0,0,0,0, 0,0,0,1,0,0,0,0,0,0,0,0), 1) # At diagnosis 2nd interval

G <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,1,0,0,0,0,0,0,0), 1) # Continuous 3rd interval
H <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,1,0,0,0,0,0,0), 1) # Discontinuous 3rd interval
I <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,1,0,0,0,0,0), 1) # Other 3rd interval
I.2 <- matrix(c(0,0,0,1,0,0,0,0,0,0,0,0,0, 0,0,0,0,0,0,0,1,0,0,0,0), 1) # At diagnosis 3rd interval

J <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,0,0,1,0,0,0), 1) # Continuous 4th interval
K <- matrix(c(0,1,0,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,0,0,0,1,0,0), 1) # Discontinuous 4th interval
L <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0,  0,0,0,0,0,0,0,0,0,0,1,0), 1) # Other 4th interval
L.2 <- matrix(c(0,0,1,0,0,0,0,0,0,0,0,0,0, 0,0,0,0,0,0,0,0,0,0,0,1), 1) # At diagnosis 4th interval

for (i in age) {
  df <- seer0to39.cp.format %>% filter(age_cat3 == i)
  x <- coxph(SurvObj ~ mcont66.2 + mcont66.2*strata(interval) + age + race_ethnicity + Yost_ACS_2006_2010_quintile + cluster(state), data = df) 
  t0<- glht(x, linfct = A) 
  df1<-confint(t0)
  df1 <- as.data.frame(exp(df1$confint))
  df1$age <- i
  df1$interval = "0-12"
  df1$category ="Continuous" # first interval
  
  t0<- glht(x, linfct = B) 
  df2<-confint(t0)
  df2 <- as.data.frame(exp(df2$confint))
  df2$age <- i
  df2$interval = "0-12"
  df2$category ="Discontinuous" # first interval
  
  t0<- glht(x, linfct = C) 
  df3<-confint(t0)
  df3 <- as.data.frame(exp(df3$confint))
  df3$age <- i
  df3$interval = "0-12"
  df3$category ="Other" # first interval
  
  t0<- glht(x, linfct = C.2) 
  df3.2<-confint(t0)
  df3.2 <- as.data.frame(exp(df3.2$confint))
  df3.2$age <- i
  df3.2$interval = "0-12"
  df3.2$category ="At diagnosis" # first interval
  
  t12<- glht(x, linfct = D) 
  df4<-confint(t12)
  df4 <- as.data.frame(exp(df4$confint))
  df4$age <- i
  df4$interval = ">12-24"
  df4$category ="Continuous" # second interval
  
  t12<- glht(x, linfct = E) 
  df5<-confint(t12)
  df5 <- as.data.frame(exp(df5$confint))
  df5$age <- i
  df5$interval = ">12-24"
  df5$category ="Discontinuous" # second interval
  
  t12<- glht(x, linfct = F) 
  df6<-confint(t12)
  df6 <- as.data.frame(exp(df6$confint))
  df6$age <- i
  df6$interval = ">12-24"
  df6$category ="Other" # second interval
  
  t12<- glht(x, linfct = F.2) 
  df6.2<-confint(t12)
  df6.2 <- as.data.frame(exp(df6.2$confint))
  df6.2$age <- i
  df6.2$interval = ">12-24"
  df6.2$category ="At diagnosis" # second interval
  
  t24<- glht(x, linfct = G) 
  df7<-confint(t24)
  df7 <- as.data.frame(exp(df7$confint))
  df7$age <- i
  df7$interval = ">24-36"
  df7$category ="Continuous" # third interval
  
  t24<- glht(x, linfct = H) 
  df8<-confint(t24)
  df8<- as.data.frame(exp(df8$confint))
  df8$age <- i
  df8$interval = ">24-36"
  df8$category ="Discontinuous" # third interval
  
  t24<- glht(x, linfct = I) 
  df9<-confint(t24)
  df9 <- as.data.frame(exp(df9$confint))
  df9$age <- i
  df9$interval = ">24-36"
  df9$category ="Other" # third interval
  
  t24<- glht(x, linfct = I.2) 
  df9.2<-confint(t24)
  df9.2 <- as.data.frame(exp(df9.2$confint))
  df9.2$age <- i
  df9.2$interval = ">24-36"
  df9.2$category ="At diagnosis" # third interval
  
  t36<- glht(x, linfct = J) 
  df10<-confint(t36)
  df10 <- as.data.frame(exp(df10$confint))
  df10$age <- i
  df10$interval = "36"
  df10$category ="Continuous" # fourth interval
  
  t36<- glht(x, linfct = K) 
  df11<-confint(t36)
  df11<- as.data.frame(exp(df11$confint))
  df11$age <- i
  df11$interval = "36"
  df11$category ="Discontinuous" # fourth interval
  
  t36<- glht(x, linfct = L) 
  df12<-confint(t36)
  df12 <- as.data.frame(exp(df12$confint))
  df12$age <- i
  df12$interval = "36"
  df12$category ="Other" # fourth interval
  
  t36<- glht(x, linfct = L.2) 
  df12.2<-confint(t36)
  df12.2 <- as.data.frame(exp(df12.2$confint))
  df12.2$age <- i
  df12.2$interval = "36"
  df12.2$category ="At diagnosis" # fourth interval
  
  
  results <-rbind(results, df1, df2, df3, df3.2, df4, df5, df6, df6.2, df7, df8, df9, df9.2, df10, df11, df12, df12.2)  # this puts the results from each loop round in the results data file
}

PHresultsmcont66.2 <- results
```

# Graph hazard ratios and confidence intervals across time for mcont66.2 variable
```{r}
PHresultsmcont66.2$interval <-  factor(PHresultsmcont66.2$interval, levels = c("0-12", ">12-24", ">24-36", "36"))
# graph HRs and 95% CIs
# 
table(PHresultsmcont66.2$interval)

# 0 to 14
p<- PHresultsmcont66.2 %>%
  filter(age == "0 to 14") %>%
  ggplot() +
  geom_point(aes(y=Estimate, x= interval, group=category, color = category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr, group=category, color = category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=16, face = "bold", color = "black"),
    axis.text.y=element_text(size=16, face = "bold", color = "black"),
    axis.title.x=element_text(size=20, face = "bold", color = "black"),
    axis.title.y=element_text(size=20, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Time interval") +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 15, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=15, face="bold")) 

p


# 15 to 39
p<- PHresultsmcont66.2 %>%
  filter(age == "15 to 39") %>%
  ggplot() +
  geom_point(aes(y=Estimate, x= interval, group=category, color = category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=Estimate, x=interval, ymax = lwr, ymin = upr, group=category, color = category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=16, face = "bold", color = "black"),
    axis.text.y=element_text(size=16, face = "bold", color = "black"),
    axis.title.x=element_text(size=20, face = "bold", color = "black"),
    axis.title.y=element_text(size=20, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Time interval") +
  theme(strip.placement = "inside", legend.position = "bottom",  strip.text = element_text(size = 15, face="bold", color="black"), strip.background = element_rect(fill="lightgray"), plot.title = element_text(face="bold"), legend.text=element_text(size=15, face="bold")) 

p
```


# SUPPLEMENTARY TABLE 7: Sensitivity analysis: exclude low grade gliomas from  Cox subtype model, OVERALL
```{r}
exclude_lowgrade <- seer0to39 %>%
  filter(lowgrade_exclude == 0)
  
cox <-function(ac, var, name){
  x <<-exclude_lowgrade %>%
    filter(age_cat3 %in% ac)
    
model<-coxph(Surv(fumo, can_dead) ~ var  + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age, and a measure of poverty

a <-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a <-a[c(1),] %>%
     rownames_to_column() %>%
    mutate(model="Adjusted for race/ethnicity, age, and Yost SES indicator") %>%
    mutate(age_group = name) %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "varEnrolled" ~ "Enrolled")) %>%
    mutate(analysis = "ex LGAST from main")
}

cox2 <-function(ac, var, name){
  x<<-exclude_lowgrade %>%
    filter(age_cat3 %in% ac)
    
model<-coxph(Surv(fumo, can_dead) ~ var +  race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age, and Yost SES index

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3,4),] %>%
    rownames_to_column() %>%
   mutate(model="Adjusted for race/ethnicity, age, and Yost SES index") %>%
    mutate(age_group = name) %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "varEnrolled" ~ "Enrolled",
                                Category == "varDiscontinuous" ~ "Discontinuous",
                                Category == "varOther" ~ "Other",
                                Category == "varContinuous" ~ "Continuous",
                                Category == "varAt diagnosis" ~ "At diagnosis")) %>%
     mutate(analysis = "ex LGAST from main")
}

a<-cox(c("0 to 14"), x$In_Medicaid_Flag, "0 to 14")
b<-cox("15 to 39", x$In_Medicaid_Flag, "15 to 39")

c<-cox2(c("0 to 14"), x$mcont66.2, "0 to 14")
d<-cox2(c("15 to 39"), x$mcont66.2, "15 to 39")

SuppTable7a<-  rbind(a,b,c,d)
```

# Astrocytomas
```{r}
exclude_lowgrade_ast <- seer0to39 %>%
  filter(lowgrade_exclude == 0, brain == "3.1 Astrocytoma") # This excludes tumors that are classifed as "3.1.1 Specified low-grade astrocytic tumors" from astrocytomas according to the website: https://seer.cancer.gov/ayarecode/aya-who2008.html
  
cox <-function(ac, var, name){
  x<<-exclude_lowgrade_ast %>%
    filter(age_cat3 %in% ac)
    
model<-coxph(Surv(fumo, can_dead) ~ var  + race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age, and a measure of poverty

a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1),] %>%
     rownames_to_column() %>%
    mutate(model="Adjusted for race/ethnicity, age, and Yost SES indicator") %>%
    mutate(age_group = name) %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "varEnrolled" ~ "Enrolled")) %>%
     mutate(analysis = "ex LGAST from AST")
}

cox2 <-function(ac, var, name){
  x<<-exclude_lowgrade_ast %>%
    filter(age_cat3 %in% ac)
    
model<-coxph(Surv(fumo, can_dead) ~ var +  race_ethnicity + age + Yost_ACS_2006_2010_quintile + cluster(state), data = x) # Model adjusted for race, age, and Yost SES index


a<-as.data.frame(exp(cbind(HR = coef(model), confint(model))))
   a<-a[c(1,2,3,4),] %>%
    rownames_to_column() %>%
   mutate(model="Adjusted for race/ethnicity, age, and Yost SES index") %>%
    mutate(age_group = name) %>%
    rename(Category = rowname) %>%
    mutate(Category = case_when(Category == "varEnrolled" ~ "Enrolled",
                                Category == "varDiscontinuous" ~ "Discontinuous",
                                Category == "varOther" ~ "Other",
                                Category == "varContinuous" ~ "Continuous",
                                Category == "varAt diagnosis" ~ "At diagnosis")) %>%
     mutate(analysis = "ex LGAST from AST")

}

a<-cox(c("0 to 14"), x$In_Medicaid_Flag, "0 to 14")
b<-cox("15 to 39", x$In_Medicaid_Flag, "15 to 39")

c<-cox2(c("0 to 14"), x$mcont66.2, "0 to 14")
d<-cox2(c("15 to 39"), x$mcont66.2, "15 to 39")

SuppTable7b<-  rbind(a,b,c,d)
```

# SUPPLEMENTARY TABLE 7

```{r}
SuppTable7 <- rbind(SuppTable7a, SuppTable7b)
```

# Clean up tables for merging for sensitivity plot comparison
```{r}
Table3_pc <- Table3 %>%
  rename(conf.low = `2.5 %`, conf.high = `97.5 %`) %>%
  select(-model) %>%
  mutate(analysis = "No exclusions")

SuppTable7_pc <- SuppTable7 %>%
  rename(conf.low = `2.5 %`, conf.high = `97.5 %`, `Age category` = age_group) %>%
  select(-model)

SuppTable4_pc <- SuppTable4 %>%
  rename(`Age category` = age_cat, Category = term, HR = estimate) %>%
  filter(btype == "3.1 Astrocytoma") %>%
  select(-c(btype, variable, abbrev )) %>%
  mutate(analysis = "No exclusions AST")

# row bind three dataframes

for_sens_plotting <- rbind(Table3_pc, SuppTable7_pc, SuppTable4_pc)

```
# Sensitivity plotting MAIN
```{r}
# 0 to 14, binary variable
p<- for_sens_plotting %>%
  filter(`Age category` == "0 to 14", Category == "Enrolled", analysis %in% c("ex LGAST from main", "No exclusions")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high, group = Category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    # axis.text.y = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "A. 0 to 14 for binary variable") +  # Add the title here
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold"))

p

library(scales)
# 0 to 14, continuity variable
q <- for_sens_plotting %>%
  filter(`Age category` == "0 to 14", Category != "Enrolled", analysis %in% c("ex LGAST from main","No exclusions")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "B. 0 to 14 for continuity variable") +  # Add the title here
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold")) +
guides(colour = guide_legend(nrow = 2))

q 

# 15 to 29, binary variable
r <- for_sens_plotting %>%
  filter(`Age category` == "15 to 39", Category == "Enrolled", analysis %in% c("ex LGAST from main", "No exclusions")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "C. 15 to 39 for binary variable") +  # Add the title here
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold"))

r

# 15 to 39, continuity variable
s <- for_sens_plotting %>%
  filter(`Age category` == "15 to 39", Category != "Enrolled", analysis %in% c("ex LGAST from main", "No exclusions")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "D. 15 to 39 for continuity variable") +  # Add the title here
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold")) +
guides(colour = guide_legend(nrow = 2))

s
```
# Sensitivity plotting AST
```{r}
table(for_sens_plotting$analysis)

# 0 to 14, binary variable
t <- for_sens_plotting %>%
  filter(`Age category` == "0 to 14", Category == "Enrolled", analysis %in% c("ex LGAST from AST", "No exclusions AST")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, group = Category, ymax = conf.low, ymin = conf.high), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "E. 0 to 14 for binary variable") +  
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold"))

t

# 0 to 19, continuous variable
u <- for_sens_plotting %>%
  filter(`Age category` == "0 to 14", Category != "Enrolled", analysis %in% c("ex LGAST from AST", "No exclusions AST")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "F. 0 to 14 for continuity variable") +  
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold"))+
guides(colour = guide_legend(nrow = 2))

u

# 15 to 39, binary variable
v <- for_sens_plotting %>%
  filter(`Age category` == "15 to 39", Category == "Enrolled", analysis %in% c("ex LGAST from AST", "No exclusions AST")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "G. 15 to 39 for binary variable") + 
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold"))

v

# 15 to 39, continuous variable
w <- for_sens_plotting %>%
  filter(`Age category` == "15 to 39", Category != "Enrolled", analysis %in% c("ex LGAST from AST", "No exclusions AST")) %>%
  ggplot() +
  geom_point(aes(y=HR, x = analysis, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), size =4) + 
  geom_hline(yintercept =1, linetype=2) +
  geom_vline(xintercept =c(1.5, 2.5,3.5,4.5,5.5,6.5,7.5,8.5,9.5), linetype=2, color = "lightgray") +
  theme(axis.text.y=element_blank()) +
  geom_errorbar(aes(y=HR, x= analysis, ymax = conf.low, ymin = conf.high, group = Category, color = Category), na.rm=TRUE, position=position_dodge(width=0.9), width = 0.1) +
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.x=element_text(size=10, face = "bold", color = "black"),
    axis.text.y=element_text(size=10, face = "bold", color = "black"),
    axis.title.x=element_text(size=10, face = "bold", color = "black"),
    axis.title.y=element_text(size=10, face = "bold", color = "black"))+
  coord_flip() +
  labs(color = NULL,  x = "Category") +
  labs(color = NULL, x = "Category", title = "H. 15 to 39 for continuity variable") +  
  theme(strip.placement = "inside", legend.position = "bottom", strip.text = element_text(size = 10, face = "bold", color = "black"), strip.background = element_rect(fill = "lightgray"), plot.title = element_text(face = "bold"), legend.text = element_text(size = 10, face = "bold"))+
guides(colour = guide_legend(nrow = 2))

w
```
# SUPPLEMENTARY FIGURE 3 and 4: Sensitivity analysis, Put all the plots together
```{r}
pdf("SUPPLEMENTARY FIGURE 3.pdf", width=11, height=8.5, paper="USr")
p + q + r + s + plot_layout(ncol=2, nrow=2) 

dev.off()

pdf("SUPPLEMENTARY FIGURE 4.pdf", width=11, height=8.5, paper="USr")
t + u + v + w + plot_layout(ncol=2, nrow=2) 

dev.off()
```
# Clean up Supplemenatary Tables
```{r}
# 1
SuppTable1 <- SuppTable1 %>%
  rename(`lower 95% CI` = lowCI,
         `upper 95% CI` = highCI,
          `Brain tumor subtype` = BrainType) %>%
  mutate_if(is.numeric, round, 2)

SuppTable1$`OR (95% CI)` <-paste0(SuppTable1$OR, " (", SuppTable1$`lower 95% CI`, " to ", SuppTable1$`upper 95% CI`, ")") 
SuppTable1 <- SuppTable1 %>%
  dplyr::select(1,6,7,8)

# 2
SuppTable2 <- SuppTable2 %>%
  rename(`lower 95% CI` = `Lower CI`,
         `upper 95% CI` = `Upper CI`)

SuppTable2$`Survival percentage  (95% CI)` <-paste0(SuppTable2$`Survival Percentage`, " (", SuppTable2$`lower 95% CI`, " to ", SuppTable2$`upper 95% CI`, ")") 

SuppTable2 <- SuppTable2 %>%
  dplyr::select(1,2,3,7)

# 3
SuppTable3 <- SuppTable3 %>%
  rename(`lower 95% CI` = LowerCI,
         `upper 95% CI` = UpperCI,
         `Age group` = Age_cat,
         `Brain tumor subtype` = Brain_tumor_subtype)


SuppTable3$`Survival percentage  (95% CI)` <-paste0(SuppTable3$Survival_Percentage, " (", SuppTable3$`lower 95% CI`, " to ", SuppTable3$`upper 95% CI`, ")") 

SuppTable3 <- SuppTable3 %>%
  dplyr::select(1,5,6,8)

# 4

SuppTable4 <- SuppTable4 %>%
  rename(`lower 95% CI` = conf.low,
         `upper 95% CI` = conf.high,
         `Age group` = age_cat,
         `Brain tumor subtype` = btype,
         HR = estimate,
         Comparison = term)%>%
  mutate_if(is.numeric, round, 2)

SuppTable4$Comparison <-paste0(SuppTable4$Comparison, " vs. Not enrolled") 

SuppTable4$`HR  (95% CI)` <-paste0(SuppTable4$HR, " (", SuppTable4$`lower 95% CI`, " to ", SuppTable4$`upper 95% CI`, ")") 

SuppTable4 <- SuppTable4 %>%
  dplyr::select(-c(2,3,4,7,8))

# 5

SuppTable5 <- SuppTable5 %>%
  rename(`lower 95% CI` = `2.5 %`,
         `upper 95% CI` = `97.5 %`,
         `Age group` = agegroup,
         Comparison = Category)%>%
  mutate_if(is.numeric, round, 2)

SuppTable5$Comparison <-paste0(SuppTable5$Comparison, " vs. Not enrolled") 

SuppTable5$`HR  (95% CI)` <-paste0(SuppTable5$HR, " (", SuppTable5$`lower 95% CI`, " to ", SuppTable5$`upper 95% CI`, ")") 

SuppTable5 <- SuppTable5 %>%
  dplyr::select(-c(2,3,4,5,8,9))

# 6

SuppTable6 <- SuppTable6 %>%
  rename(`lower 95% CI` = `lwr`,
         `upper 95% CI` = `upr`,
         HR = Estimate,
         Comparison = category, 
         `Age group/brain tumor subtype` = age_can) %>%
         mutate_if(is.numeric, round, 2)

SuppTable6$Comparison <-paste0(SuppTable6$Comparison, " vs. Not enrolled") 

SuppTable6$`HR  (95% CI)` <-paste0(SuppTable6$HR, " (", SuppTable6$`lower 95% CI`, " to ", SuppTable6$`upper 95% CI`, ")") 

SuppTable6 <- SuppTable6 %>%
  dplyr::select(4,5,6,7) %>%
  dplyr::select(3,1,2,4)

# 7

SuppTable7 <- SuppTable7 %>%
  rename(`lower 95% CI` = `2.5 %`,
         `upper 95% CI` = `97.5 %`,
         Comparison = Category, 
         `Age group` = age_group) %>%
         mutate_if(is.numeric, round, 2)

SuppTable7$Comparison <-paste0(SuppTable7$Comparison, " vs. Not enrolled") 

SuppTable7$`HR  (95% CI)` <-paste0(SuppTable7$HR, " (", SuppTable7$`lower 95% CI`, " to ", SuppTable7$`upper 95% CI`, ")") 

SuppTable7 <- SuppTable7 %>%
  dplyr::select(-c(2,3,4,5)) 
```

# Export supplementary tables to excel
```{r}
sheets <- list("Table 2" = Table2,
               "Table 3" = Table3)

write.xlsx(sheets, file="Tables.xlsx", showNA=TRUE, overwrite = TRUE)

sheets <- list("Supp Table 1 (stage btype)" = SuppTable1,
               "Supp Table 2 (surv 60 mo)" = SuppTable2,
               "Supp Table 3 (surv 60 btype)" = SuppTable3,
               "Supp Table 4 (HRs by btype)" = SuppTable4,
               "Supp Table 5 (HRs by dem)" = SuppTable5,
               "Supp Table 6 (HRs by time)" = SuppTable6)
write.xlsx(sheets, file="Supplementary tables.xlsx", showNA=TRUE, overwrite = TRUE)
```

# try propensity score matching for In_Medicaid_Flag
```{r}
# Install MatchIt
# install.packages("MatchIt")
library(MatchIt)

matched <- matchit(In_Medicaid_Flag ~ race_ethnicity + Yost_ACS_2006_2010_quintile + sex + age + brain , data = seer0to39, distance = "glm", method = "nearest", ratio = 2)

summary(matched)
plot(matched, type = "jitter")

m.data <- match.data(matched)

# adjusted models 0 to 39 (original model)
res.cox2 <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + cluster(state), data = seer0to39) 
res.cox2
summary(res.cox2)

# adjusted models 0 to 39 (with propensity score matching)
res.cox2p <- coxph(Surv(fumo, can_dead) ~ In_Medicaid_Flag + cluster(state), weights = m.data$weights, data = m.data) 
res.cox2p
summary(res.cox2p)
```

# example for ph
```{r}
fit1 <- coxph(Surv(time, status) ~ karno + age + trt, veteran)
plot(cox.zph(fit1)[1])
# a cox.zph plot of the data suggests that the effect of Karnofsky score
# begins to diminish by 60 days and has faded away by 120 days.
# Fit a model with separate coefficients for the three intervals.

veteran <- veteran %>%
  mutate(ID = row_number())%>%
  mutate(karno2 = case_when(karno <60 ~0,
                            karno >=60 ~1))

vet2 <- survSplit(Surv(time, status) ~., veteran,
                   cut=c(60, 120), 
                    end   = "time",    # original survival time
                    event = "status",   # event indicator
                    episode ="timegroup")

# Recoding
vet2 <- within(vet2, {
    # Create new survival object
    SurvObj <- Surv(tstart, time, status) # creates column labeled col
    # Create interval indicator
    interval <- factor(tstart, levels = c(0,60,120), labels = c("First","Second", "Third"))
})

# Reordering
vet2 <- vet2[with(vet2, order(ID, time)),]  # sort by id and then by time


# Fit extended Cox model

# Testing the rx variable only
# cluster(id) for robust SE to account for within-cluster non-independence

# results match from two approaches when only exposure of interest and time variable in the model, diverge more when adjustment variables included and diverge less when these are wrapped by strata. Not sure what approach is best but will use Kleinman suggestion of heavyside approach and include interaction terms in model
fit2 <- coxph(SurvObj ~ karno2 + karno2*strata(timegroup) + age + cluster(ID), data= vet2)
summary(fit2)

# cov <- vcov(fit2)
# time intervals using contrasts
library(multcomp)

X<-matrix(c(1, 0, 1, 0), 1)

t <- glht(fit2, linfct = X) 
summary(t) # here are the results

CI<-confint(t)
CI

fit3 <- coxph(SurvObj ~ karno2 + age +  cluster(ID), data= vet2[which(vet2$timegroup == 3),])

summary(fit3)
```


# EXPLORATORY ANALYSIS: TREATMENT DIFFERENCES RADIATION AND SURGERY 
- logistic analysis to see if Medicaid predicts radiation and surgery
- 0 to 19
```{r}
# Surgery In_Medicaid_Flag
model1 <- miceadds::glm.cluster(seer0to39 %>% filter(age_cat2 == "0 to 19", surgery != "Unknown"), surgery ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile + stage, cluster="state", family="binomial")
summary(model1)

# The odds of surgery was 22% lower in those with Medicaid

# Radiation In_Medicaid_Flag
model2 <- miceadds::glm.cluster(seer0to39_ex %>% filter(age_cat2 == "0 to 19", radiation != "Unknown", surgery == "No surgery"), radiation ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile + stage, cluster="state", family="binomial")
summary(model2)

# The odds of radiation was 1.64 times higher in those with Medicaid

# exclude low grades

# Surgery In_Medicaid_Flag
model1a <- miceadds::glm.cluster(seer0to39 %>% filter(age_cat2 == "0 to 19", surgery != "Unknown", lowgrade_exclude == 0), surgery ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model1a)

# Radiation In_Medicaid_Flag
model2a <- miceadds::glm.cluster(seer0to39 %>% filter(age_cat2 == "0 to 19", radiation != "Unknown", lowgrade_exclude == 0), radiation ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model2a)

# Surgery mcont66.2
model3 <- miceadds::glm.cluster(seer0to39 %>% filter(age_cat2 == "0 to 19", surgery != "Unknown"), surgery ~ mcont66.2 + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model3)

# Radiation mcont66.2
model4 <- miceadds::glm.cluster(seer0to39 %>% filter(age_cat2 == "0 to 19", radiation != "Unknown"), radiation ~ mcont66.2 + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model4)
```

# 20 to 39
```{r}
# Surgery In_Medicaid_Flag
model1 <- miceadds::glm.cluster(seer0to39_ex %>% filter(age_cat2 == "20 to 39", surgery != "Unknown"), surgery ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model1)

# The odds of surgery was 18% lower in those with Medicaid

# Radiation In_Medicaid_Flag
model2 <- miceadds::glm.cluster(seer0to39_ex %>% filter(age_cat2 == "20 to 39", radiation != "Unknown"), radiation ~ In_Medicaid_Flag + race_cat2 + age + brain2 + Yost_ACS_2006_2010_quintile, cluster="state", family="binomial")
summary(model2)

# The odds of radiation was 1.43 times higher in those with Medicaid
```

